---
title: "MES-like NB tumor cells with Sören S."
author: "Sara Wernig-Zorc"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
    collapsed: no
    code_folding: hide
    highlight: pygments
    theme: spacelab
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '5'
params:
  input: "/home/rstudio/mnt_data/"
  out: "/home/rstudio/mnt_out/"
  data: "/home/rstudio/mnt_data/"
  nb_tumors: "/home/rstudio/mnt_nb_tumors/"
  GO_db: "/home/rstudio/mnt_resources/GSEA_GeneSetEnrichments/"
  markers: "/home/rstudio/mnt_resources/MES_ADRN/"
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      tidy = TRUE, 
                      error = FALSE, 
                      messages = FALSE,
                      warning = FALSE)

library(parallel)
future::plan(strategy = 'future::multicore', workers = 20)
options(future.globals.maxSize = 8 * 10^9)

print(params)


source("/home/rstudio/scRNA.scATAC/Rmd/MES_project/MES_project_setup.R")
```

```{r R packages: load}
#install.packages("Matrix", repos = "http://R-Forge.R-project.org")
suppressPackageStartupMessages(library("Seurat"))
suppressPackageStartupMessages(library("SeuratWrappers"))
suppressPackageStartupMessages(library("SeuratData"))
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("scCustomize"))
suppressPackageStartupMessages(library("RColorBrewer"))
suppressPackageStartupMessages(library("SCpubr"))
suppressPackageStartupMessages(library("ggVennDiagram"))
suppressPackageStartupMessages(library("statmod"))
suppressPackageStartupMessages(library("miloR"))
suppressPackageStartupMessages(library("SingleCellExperiment"))
suppressPackageStartupMessages(library("scater"))
suppressPackageStartupMessages(library("scran"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("patchwork"))
```

# __Patel et al., 2024__

The dataset contains scRNA-seq (n=13 tumors; 84,769 cells) and 
snRNA-seq (n=43 tumors; 445,286 nuclei) samples
<br>
In the pre-print Patel et al., 2024 they show clusters 4, 6 and 11 as MES,
cluster 1 as SYM and the rest as ADRN

```{r eval=FALSE, include=FALSE}
# loading the Robject with only tumor cells 
# defined by the authors Patel et al. based on inferCNV
patel2024 <- load(file = paste0(params$nb_tumors,                             "/patel2024.scRNA/data/scRNA-seq/Rds/2020_08_18_all_Malignant_batchcorrect_liger_k12.Rda"))

gcdata@meta.data
  # annotate_refine_1
  # Age
  # Gender
  # MYCN
  # ALK
  # INSS_Stage
  # INRGSS_Stage
  # Treatment_Status

table(gcdata@meta.data$SingleR.cell.labels)

gcdata@meta.data$sample_name <- 
  gcdata@meta.data$orig.ident

saveRDS(gcdata, file = paste0(params$nb_tumors,
                                    "/patel2024.scRNA/data/scRNA-seq/Rds/patel2024_tumor_cluster.Rds"))
```

```{r}
tumorCells <- readRDS(file = paste0(params$nb_tumors,
                                    "/patel2024.scRNA/data/scRNA-seq/Rds/patel2024_tumor_cluster.Rds"))
```

```{r eval=FALSE, include=FALSE}
tumorCells@meta.data$sample_name <- 
  tumorCells@meta.data$orig.ident

metadata <- read.csv(file = paste0(params$nb_tumors,
                                   "/patel2024.scRNA/data/scRNA-seq/metadata/metadata.csv"),
                     sep = "\t", 
                     header = TRUE)

table(tumorCells@meta.data$Histologic_Diagnosis, 
      tumorCells@meta.data$sample_name)

Idents(tumorCells) <- tumorCells@meta.data$liger_clusters

# metadata$sample_name == highrisk_patients@meta.data$sample_name

old_metadata <- as.data.frame(tumorCells@meta.data)

merged_metadata <- merge(
  old_metadata, 
  metadata, 
  by = "sample_name", 
  all.x = TRUE
)

rownames(merged_metadata) <- rownames(tumorCells@meta.data)  
tumorCells@meta.data <- merged_metadata

table(tumorCells@meta.data$treatment)
```


```{r}
filter_by_risk <- tumorCells$COG_Risk == "High"

highrisk_patients <- tumorCells[,filter_by_risk]
table(highrisk_patients$sample_name)
table(highrisk_patients$Histologic_Diagnosis)

filter_by_histology <- highrisk_patients$Histologic_Diagnosis  != "GNB"

NB_HR_patients <- highrisk_patients[,filter_by_histology]
table(NB_HR_patients$orig.ident)

filter_by_treatment <- NB_HR_patients$Treatment_Status  != "Unknown"
NB_HR_patients_complete <- NB_HR_patients[,filter_by_treatment]

table(NB_HR_patients_complete$Treatment_Status,NB_HR_patients_complete$orig.ident)

# 20 patients remaining

DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST_v2, 
        raster = FALSE,
        shuffle = TRUE,
        label.size = 36)

table(NB_HR_patients_complete$liger_clusters)

saveRDS(NB_HR_patients_complete, 
        file = paste0(params$out,
                      "patel2024/Rds/patel2024_NB-HR_tumor_cluster.Rds"))
```

```{r}
# 20 patients with NB histology, high-risk tumors and information on treatment
NB_HR_patients_complete <- readRDS(file = paste0(params$out,
                                  "patel2024/Rds/patel2024_NB-HR_tumor_cluster.Rds"))
```

# ___MES/ADR gene list___

```{r fig.height=5, fig.width=5}
genes_of_interest <- c("YAP1","WWTR1","JUN","FOSL1","FOSL2")
dotplot_genes <- c("FN1","VIM","CD44",
                   "PHOX2B","DBH",
                   "SOX10","S100B",
                   "YAP1","WWTR1",
                   "JUN","FOSL1","FOSL2")
dotplot_genes_v2 <- c("FN1","VIM","CD44",
                   "PHOX2B","DBH",
                   "YAP1","WWTR1",
                   "JUN","FOSL1","FOSL2")
dotplot_genes_v3 <- c("FN1","VIM","CD44",
                   "PHOX2B","DBH",
                   "SOX10","S100B","PLP1",
                   "YAP1","WWTR1",
                   "JUN","FOSL1","FOSL2")

dotplot_genes_v4 <- c("FN1","VIM","CD44",
                   "PHOX2B","DBH",
                   "SOX10","PLP1",
                   "YAP1","WWTR1",
                   "JUN","FOSL1","FOSL2")


core_ADR_markers <-  c("PHOX2B","DBH","GATA3","TH")
core_MES_markers <- c("CD44","VIM","FN1","COL4A1")
core_SYM_markers <- c("MKI67","TOP2A")

# Typical MES/ADR markers
MES_markers <- 
  read.table(file = paste0(params$markers,"/MES_markers.txt"),
             sep = "\t", header=FALSE, row.names=1)
MES_markers <- MES_markers$V2

ADR_markers <- 
  read.table(file = paste0(params$markers,"/ADRN_markers.txt"),
             sep = "\t", header=FALSE, row.names=1)
ADR_markers <- ADR_markers$V2

# MES/ADR markers from Thirant et al., 2023 (from NB patients)
thirant2023_MES_genes <- 
  read.table(file = paste0(params$markers,"thirant2023_MES_all_genes_patients.txt"),
             sep = "\t", header=FALSE, quote = "")
thirant2023_MES_genes <- as.list(thirant2023_MES_genes)

thirant2023_ADR_SYM_genes <- 
  read.table(file = paste0(params$markers,"thirant2023_ADR_SYM_genes_patients.txt"),
             sep = "\t", header=FALSE)
thirant2023_ADR_SYM_genes <- as.list(thirant2023_ADR_SYM_genes)

thirant2023_ADR_CHR_genes <- 
  read.table(file = paste0(params$markers,"thirant2023_ADR_CHRO_genes_patients.txt"),
             sep = "\t", header=FALSE)
thirant2023_ADR_CHR_genes <- as.list(thirant2023_ADR_CHR_genes)

# MES/ADR markers from cell lines (bulk RNA-seq data from Sören)
soren2024_MES_genes <- 
  read.table(file = paste0(params$markers,"DEGs_NB_cell_lines_MES_specific_genes.txt"),
             sep = "\t", header=FALSE)
soren2024_MES_genes <- as.list(soren2024_MES_genes)

soren2024_ADR_genes <- 
  read.table(file = paste0(params$markers,"DEGs_NB_cell_lines_ADR_specific_genes.txt"),
             sep = "\t", header=FALSE)
soren2024_ADR_genes <- as.list(soren2024_ADR_genes)

# MES/ADR markers from cell lines (bulk RNA-seq data from Sören)
# fold change above 3.3 (-3.3 for ADR genes)
soren2024_MES_genes_v2 <- 
  read.table(file = paste0(params$markers,"strohmenger2024_MES_markers_from_NB_cell_lines_foldchange+10.txt"),
             sep = "\t", header=FALSE)
soren2024_MES_genes_v2 <- as.list(soren2024_MES_genes_v2)

soren2024_ADR_genes_v2 <- 
  read.table(file = paste0(params$markers,"strohmenger2024_ADR_markers_from_NB_cell_lines_foldchange-10.txt"),
             sep = "\t", header=FALSE)
soren2024_ADR_genes_v2 <- as.list(soren2024_ADR_genes_v2)

# MES/ADR markers from cell lines (bulk RNA-seq data from Sören)
# log2 fold change above 5 (-5 for ADR genes)

soren2024_MES_genes_v3 <- 
  read.table(file = paste0(params$markers,"strohmenger2024_MES_markers_from_NB_cell_lines_log2FC+5.txt"),
             sep = "\t", header=FALSE)
soren2024_MES_genes_v3 <- as.list(soren2024_MES_genes_v3)

soren2024_ADR_genes_v3 <- 
  read.table(file = paste0(params$markers,"strohmenger2024_ADR_markers_from_NB_cell_lines_log2FC-5.txt"),
             sep = "\t", header=FALSE)
soren2024_ADR_genes_v3 <- as.list(soren2024_ADR_genes_v3)


# Gene list from Boeva et al., 2021
boeva2021_ADR_genes <- c("PHOX2B","GATA3","HAND2","HAND1","KLF7","ISL1","PHOX2A")
                  
boeva2021_NCC_genes <- c("NR3C1","BHLHE41","MAFF","GLIS3","IRF1","IRF2","IRF3",
                        "FLI1","MEF2D","PRRX1","RUNX1","RUNX2","TBX18","FOSL1",
                        "FOSL2")

# Gene list from Van Groeningen et al., 2017
vangroningen2017_ADR_genes <- 
  read.table(paste0(params$markers,"vangroningen2017_ADR_genes.txt"),
                          header = FALSE, quote = "")
vangroningen2017_ADR_genes <- as.list(vangroningen2017_ADR_genes)
vangroningen2017_MES_genes <- 
  read.csv(paste0(params$markers,"vangroningen2017_MES_genes.txt"),
                          header = FALSE, quote = "")
vangroningen2017_MES_genes <- as.list(vangroningen2017_MES_genes)

gene.lists <- list("genes_of_interest" = genes_of_interest,
                "core_ADR_markers" = core_ADR_markers,
                "core_MES_markers" = core_MES_markers,
                "core_SYM_markers" = core_SYM_markers,
                "MES_markers" = MES_markers,
                "ADR_markers" = ADR_markers,
                "thirant2023_MES_genes" = thirant2023_MES_genes$V1,
                "thirant2023_ADR_SYM_genes" = thirant2023_ADR_SYM_genes$V1,
                "thirant2023_ADR_CHR_genes" = thirant2023_ADR_CHR_genes$V1,
                "soren2024_MES_genes" = soren2024_MES_genes$V1,
                "soren2024_ADR_genes" = soren2024_ADR_genes$V1,
                "boeva2021_ADR_genes" = boeva2021_ADR_genes,
                "boeva2021_NCC_genes" = boeva2021_NCC_genes,
                "vangroningen2017_ADR_genes" = vangroningen2017_ADR_genes$V1,
                "vangroningen2017_MES_genes" = vangroningen2017_MES_genes$V1)
```

## ___Gene list overlaps: all DEGs___

```{r Venn diagram ADR fig.height=5, fig.width=10}
# overlap between published ADR gene lists 
# and the new gene list in this publication
ADR_gene_lists <- list(vangroningen2017 = vangroningen2017_ADR_genes$V1, 
                       strohmenger2024 = soren2024_ADR_genes$V1,
                       boeva2021 = boeva2021_ADR_genes)

#pdf(file = paste0(params$out,"NB_HR_patients_complete/figures/ADR_gene_lists_overlap.pdf"),
#    height = 5, width = 10)
ggVennDiagram(ADR_gene_lists, set_size = 6, label_alpha = 0,label_size = 6,
  category.names = c("Van Groningen","Strohmenger","Boeva"),
  edge_size = 1, label_percent_digit = 1,label_color = "white") +
  ggplot2::scale_fill_gradient(low="#B59410",high = "darkblue") +
  labs(title = "\t \t \t \t \t \t \t \t \t Overlap of ADR gene lists") + 
  theme(title = element_text(size = 16),legend.text = element_text(size = 14)) +
  scale_x_continuous(expand = expansion(mult = .2))
#dev.off()

ADR_gene_lists_v2 <- list(vangroningen2017 = vangroningen2017_ADR_genes$V1, 
                       strohmenger2024 = soren2024_ADR_genes$V1,
                       boeva2021 = boeva2021_ADR_genes,
                       thirant2023a = thirant2023_ADR_SYM_genes$V1,
                       thirant2023b = thirant2023_ADR_CHR_genes$V1)


ggVennDiagram(ADR_gene_lists_v2,  set_size = 8, 
              label_alpha = 0, label_size = 8,
  category.names = c("vangroningen2017","strohmenger2024","boeva2021",
                     "thirant2023a","thirant2023b")) +
  ggplot2::scale_fill_gradient(low="#B59410",high = "darkblue")  +
  labs(title = "\t \t \t \t Overlap of ADR gene lists") + 
  theme(title = element_text(size = 20),legend.text = element_text(size = 18)) +
  scale_x_continuous(expand = expansion(mult = .4))
```

```{r Venn diagram MES fig.height=5, fig.width=10}
# overlap between published MES gene lists 
# and the new gene list in this publication
MES_gene_lists <- list(vangroningen2017 = vangroningen2017_MES_genes$V1, 
                       strohmenger2024 = soren2024_MES_genes$V1,
                       boeva2021 = boeva2021_NCC_genes)

#pdf(file = paste0(params$out,"NB_HR_patients_complete/figures/MES_gene_lists_overlap.pdf"),
#    height = 5, width = 10)
ggVennDiagram(MES_gene_lists, set_size = 6, label_alpha = 0,label_size = 6,
  category.names = c("Van Groningen","Strohmenger","Boeva"),
  edge_size = 1, label_percent_digit = 1,label_color = "white") +
  ggplot2::scale_fill_gradient(low="#B59410",high = "darkblue") +
  labs(title = "\t \t \t \t \t \t \t \t \t Overlap of MES gene lists") + 
  theme(title = element_text(size = 16),legend.text = element_text(size = 14)) +
  scale_x_continuous(expand = expansion(mult = .2))
#dev.off()

MES_gene_lists_v2 <- list(vangroningen2017 = vangroningen2017_MES_genes$V1, 
                       strohmenger2024 = soren2024_MES_genes$V1,
                       boeva2021 = boeva2021_NCC_genes,
                       thirant2023a = thirant2023_MES_genes$V1)


ggVennDiagram(MES_gene_lists_v2,  set_size = 8, 
              label_alpha = 0, label_size = 8,
  category.names = c("vangroningen2017","strohmenger2024","boeva2021",
                     "thirant2023")) +
  ggplot2::scale_fill_gradient(low="#B59410",high = "darkblue")  +
  labs(title = "\t \t \t \t Overlap of MES gene lists") + 
  theme(title = element_text(size = 20),legend.text = element_text(size = 18)) +
  scale_x_continuous(expand = expansion(mult = .4))
```

## ___Gene list overlaps: top DEGs___

```{r set cut-off for DEGs}
cell_type_MES_vs_ADR <- 
  read_excel(paste0(params$markers,"cell_type_MES_vs_ADR.xlsx"))

# setting a cut-off for DEGs
##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/DEGs_log2FC_distribution_histogram.png"))
hist(cell_type_MES_vs_ADR$log2FoldChange,
     plot = TRUE,
     n=100,
     col = "darkgreen", 
     main = "Distribution of log2 fold changes (significant)")
abline(v = -2,col="lightblue",lwd=2)
abline(v = 2,col="lightblue",lwd=2)
abline(v = -3.3,col="darkblue",lwd=2)
abline(v = 3.3,col="darkblue",lwd=2)
abline(v = -5,col="purple",lwd=2)
abline(v = 5,col="purple",lwd=2)
#dev.off()
```

```{r Venn diagram ADR fig.height=5, fig.width=10}
# overlap between published ADR gene lists 
# and the new gene list in this publication
ADR_gene_lists <- list(vangroningen2017 = vangroningen2017_ADR_genes$V1, 
                       strohmenger2024 = soren2024_ADR_genes_v2$V1,
                       boeva2021 = boeva2021_ADR_genes)

#pdf(file = paste0(params$out,"NB_HR_patients_complete/figures/ADR_gene_lists_overlap.pdf"),
#    height = 5, width = 10)
ggVennDiagram(ADR_gene_lists, set_size = 8, label_alpha = 0,label_size = 8,
  category.names = c("Van Groningen","Strohmenger","Boeva"),
  edge_size = 1, label_percent_digit = 1,label_color = "white") +
  ggplot2::scale_fill_gradient(low="#B59410",high = "darkblue") +
  labs(title = "\t \t \t \t \t \t \t Overlap of ADR gene lists") + 
  theme(title = element_text(size = 20),legend.text = element_text(size = 18)) +
  scale_x_continuous(expand = expansion(mult = .3))
#dev.off()

ADR_gene_lists_v2 <- list(vangroningen2017 = vangroningen2017_ADR_genes$V1, 
                       strohmenger2024 = soren2024_ADR_genes_v2$V1,
                       boeva2021 = boeva2021_ADR_genes,
                       thirant2023a = thirant2023_ADR_SYM_genes$V1,
                       thirant2023b = thirant2023_ADR_CHR_genes$V1)


ggVennDiagram(ADR_gene_lists_v2,  set_size = 8, 
              label_alpha = 0, label_size = 8,
  category.names = c("vangroningen2017","strohmenger2024","boeva2021",
                     "thirant2023a","thirant2023b")) +
  ggplot2::scale_fill_gradient(low="#B59410",high = "darkblue")  +
  labs(title = "\t \t \t \t Overlap of ADR gene lists") + 
  theme(title = element_text(size = 20),legend.text = element_text(size = 18)) +
  scale_x_continuous(expand = expansion(mult = .4))
```

```{r Venn diagram MES fig.height=5, fig.width=10}
# overlap between published MES gene lists 
# and the new gene list in this publication
MES_gene_lists <- list(vangroningen2017 = vangroningen2017_MES_genes$V1, 
                       strohmenger2024 = soren2024_MES_genes_v2$V1,
                       boeva2021 = boeva2021_NCC_genes)

#pdf(file = paste0(params$out,"NB_HR_patients_complete/figures/MES_gene_lists_overlap.pdf"),
#    height = 5, width = 10)
ggVennDiagram(MES_gene_lists, set_size = 8, label_alpha = 0,label_size = 8,
  category.names = c("Van Groningen","Strohmenger","Boeva"),
  edge_size = 1, label_percent_digit = 1,label_color = "white") +
  ggplot2::scale_fill_gradient(low="#B59410",high = "darkblue") +
  labs(title = "\t \t \t \t \t \t \t Overlap of MES gene lists") + 
  theme(title = element_text(size = 20),legend.text = element_text(size = 18)) +
  scale_x_continuous(expand = expansion(mult = .3))
#dev.off()

MES_gene_lists_v2 <- list(vangroningen2017 = vangroningen2017_MES_genes$V1, 
                       strohmenger2024 = soren2024_MES_genes_v2$V1,
                       boeva2021 = boeva2021_NCC_genes,
                       thirant2023a = thirant2023_MES_genes$V1)


ggVennDiagram(MES_gene_lists_v2,  set_size = 8, 
              label_alpha = 0, label_size = 8,
  category.names = c("vangroningen2017","strohmenger2024","boeva2021",
                     "thirant2023")) +
  ggplot2::scale_fill_gradient(low="#B59410",high = "darkblue")  +
  labs(title = "\t \t \t \t Overlap of MES gene lists") + 
  theme(title = element_text(size = 20),legend.text = element_text(size = 18)) +
  scale_x_continuous(expand = expansion(mult = .4))
```



## ___ADR/MES module scores___

```{r filter gene lists}
# extract counts
counts <- LayerData(NB_HR_patients_complete, assay="RNA", layer='data')
var_genes <- dimnames(counts)[[1]]

# remove genes not present in dataset
soren2024_MES_genes <- intersect(soren2024_MES_genes$V1, var_genes)
soren2024_ADR_genes <- intersect(soren2024_ADR_genes$V1, var_genes)
MES_markers <- intersect(MES_markers, var_genes)
ADR_markers <- intersect(ADR_markers, var_genes)

thirant2023_MES_genes <- intersect(thirant2023_MES_genes$V1, var_genes)
thirant2023_ADR_SYM_genes <- intersect(thirant2023_ADR_SYM_genes$V1, var_genes)
thirant2023_ADR_CHR_genes <- intersect(thirant2023_ADR_CHR_genes$V1, var_genes)

vangroningen2017_ADR_genes <- intersect(vangroningen2017_ADR_genes$V1, var_genes)
vangroningen2017_MES_genes <- intersect(vangroningen2017_MES_genes$V1, var_genes)

soren2024_MES_genes_v2 <- intersect(soren2024_MES_genes_v2$V1, var_genes)
soren2024_ADR_genes_v2 <- intersect(soren2024_ADR_genes_v2$V1, var_genes)

soren2024_ADR_genes_v3 <- intersect(soren2024_ADR_genes_v3$V1, var_genes)
soren2024_MES_genes_v3 <- intersect(soren2024_MES_genes_v3$V1, var_genes)

boeva2021_ADR_genes <- intersect(boeva2021_ADR_genes, var_genes)
boeva2021_NCC_genes <- intersect(boeva2021_NCC_genes, var_genes)
```

```{r eval=FALSE, include=FALSE}
# loop through different lists and add module score to the Seurat object
for (name in names(gene.lists)) {
  print(paste0("Processing list: ", name))
  for (i in 1:length(gene.lists)) {
    #print(gene.lists[i])
    NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                             features = gene.lists[i],
                             #ctrl = 25,
                             name = name,
                             seed = 28)
  }
}
```

```{r generate moduleScore}
NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = genes_of_interest,
                         name = "genes_of_interest",
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = core_ADR_markers,
                         name = "core_ADR_markers",
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = core_MES_markers,
                         name = "core_MES_markers",
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = core_SYM_markers,
                         name = "core_SYM_markers",
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = MES_markers,
                         name = "MES_markers",         
                         seed = 28)
NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = ADR_markers,
                         name = "ADR_markers",           
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = thirant2023_MES_genes,
                         name = "thirant2023_MES_genes",
                         seed = 28)
NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = thirant2023_ADR_SYM_genes,
                         name = "thirant2023_ADR_SYM_genes",
                         seed = 28)           
#NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
#                         features = thirant2023_ADR_CHR_genes,
#                         name = "thirant2023_ADR_CHR_genes",
#                         seed = 28)       #Error in features[[i]] : subscript out of bounds     

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = soren2024_MES_genes,
                         name = "soren2024_MES_genes",
                         seed = 28)
NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = soren2024_ADR_genes,
                         name = "soren2024_ADR_genes",
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = soren2024_MES_genes_v2,
                         name = "soren2024_MES_genes_v2",
                         seed = 28)
NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = soren2024_ADR_genes_v2,
                         name = "soren2024_ADR_genes_v2",
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = soren2024_MES_genes_v3,
                         name = "soren2024_MES_genes_v3",
                         seed = 28)
NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = soren2024_ADR_genes_v3,
                         name = "soren2024_ADR_genes_v3",
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = boeva2021_NCC_genes,
                         name = "boeva2021_NCC_genes",
                         seed = 28)
NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = boeva2021_ADR_genes,
                         name = "boeva2021_ADR_genes",
                         seed = 28)

NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = vangroningen2017_ADR_genes,
                         name = "vangroningen2017_ADR_genes",
                         seed = 28)
NB_HR_patients_complete <- AddModuleScore(NB_HR_patients_complete,
                         features = vangroningen2017_MES_genes,
                         name = "vangroningen2017_MES_genes",
                         seed = 28)
```

## ___ADR/MES Module score plots___

```{r plot moduleScores: thirant}

#### Thirant et al., 2023 ####

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_thirant2023_MES_genes.pdf"),
#    height = 5, width = 5)
FeaturePlot(NB_HR_patients_complete, 
            features = "thirant2023_MES_genes1", 
            col = c("gray","orange","red"),
            order=TRUE,
            pt.size = 2, 
            label = FALSE,
            label.size = 12)          

#dev.off()

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_thirant2023_ADR_SYM_genes.pdf"), 
#    height = 5, width = 5)
FeaturePlot(NB_HR_patients_complete, 
            features = "thirant2023_ADR_SYM_genes1", 
            col = c("gray","orange","red"),
            order=TRUE,
            pt.size = 2, 
            label = FALSE,
            label.size = 12)
#dev.off()
```

```{r moduleScores: Boeva, fig.height=5, fig.width=4}

#### Boeva et al., 2021 #####

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_Boeva2021_NCC_genes.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "boeva2021_NCC_genes1",
               plot.title = "NCC gene module score",
               plot.subtitle = "Boeva et al., 2021",
               order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression")
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_Boeva2021_NCC_genes.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "boeva2021_NCC_genes1",
               plot.title = "NCC gene module score",
               plot.subtitle = "Boeva et al., 2021",
               order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression")
#dev.off()

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_Boeva2021_ADR_genes.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "boeva2021_ADR_genes1",
               plot.title = "ADR gene module score",
               plot.subtitle = "Boeva et al., 2021",
               order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression")
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_Boeva2021_ADR_genes.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "boeva2021_ADR_genes1",
               plot.title = "ADR gene module score",
               plot.subtitle = "Boeva et al., 2021",
               order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression")
#dev.off()
```

```{r plot moduleScores: Van Groeningen, fig.height=5, fig.width=4}
#### Van Groeningen et al., 2017 ####
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_vangroningen2017_MES_genes_shuffle.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "vangroningen2017_ADR_genes1",
               plot.title = "ADR gene module score",
               plot.subtitle = "Van Groningen et al., 2017",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression")
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_vangroningen2017_MES_genes.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "vangroningen2017_ADR_genes1",
               plot.title = "ADR gene module score",
               plot.subtitle = "Van Groningen et al., 2017",
               order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression")
#dev.off()

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_vangroningen2017_ADR_genes_shuffle.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "vangroningen2017_MES_genes1",
               plot.title = "MES gene module score",
               plot.subtitle = "Van Groningen et al., 2017",
               viridis.direction = -1,
               viridis.palette = "D",
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression")
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/ModuleScore_vangroningen2017_ADR_genes.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "vangroningen2017_MES_genes1",
               plot.title = "MES gene module score",
               plot.subtitle = "Van Groningen et al., 2017",
               order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression")
#dev.off()
```

```{r plot moduleScores: Sören-defined-markers, fig.height=5, fig.width=4}
# using version 1 gene list
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "soren2024_MES_genes1",
               plot.title = "soren2024 MES genes",
               order = TRUE,
               pt.size = 0.25,
               viridis.palette = "D",
               viridis.direction = -1,
               font.size = 14,
               label.size = 14,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)

do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "soren2024_ADR_genes1",
               plot.title = "soren2024 ADR genes",
               order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 14,
               label.size = 14,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)

# Using version 2 gene list
#pdf(file = paste0(params$out, "/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_Soren2024_MES_markers.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "soren2024_MES_genes_v21",
               plot.title = "soren2024 MES genes",
               plot.subtitle = "DEGs with fold change > 10",
            #   raster = TRUE,
               order = TRUE,
               pt.size = 0.25,
               viridis.palette = "D",
               viridis.direction = -1,
               font.size = 18,
               label.size = 18,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_Soren2024_MES_markers_FINAL.png"), 
#    height = 5900, width = 6470, res = 400)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "soren2024_MES_genes_v21",
               plot.title = "\t\t MES (this study)",
               order = TRUE,
               pt.size = 2,
               viridis.palette = "D",
               viridis.direction = -1,
               font.size = 60,
               label.size = 55,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "",
               legend.width = 4,
               legend.length = 20,
               legend.position = "right")
#dev.off()

#pdf(file = paste0(params$out, "/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_Soren2024_ADR_markers.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "soren2024_ADR_genes_v21",
               plot.title = "soren2024 ADR genes",
               plot.subtitle = "DEGs with fold change < -10",
                raster = TRUE,
               order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 14,
               label.size = 14,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_Soren2024_ADR_markers_FINAL.png"), 
#    height = 5900, width = 6470, res = 400)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "soren2024_ADR_genes_v21",
               plot.title = "\t\t ADR (this study)",
               order = TRUE,
               pt.size = 2,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 60,
               label.size = 55,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "",
               legend.width = 4,
               legend.length = 20,
               legend.position = "right")
#dev.off()

# Using version 3 gene list

do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "soren2024_MES_genes_v31",
               plot.title = "soren2024 MES genes",
               order = TRUE,
               pt.size = 0.25,
               viridis.palette = "D",
               viridis.direction = -1,
               font.size = 14,
               label.size = 14,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)

do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "soren2024_ADR_genes_v31",
               plot.title = "soren2024 ADR genes",
               order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 14,
               label.size = 14,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
```

```{r plot genes of interest}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_genes_of_interest.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "genes_of_interest1",
               plot.title = "MES CRC factors",
               plot.subtitle = "(YAP1, WWTR1, JUN, FOSL1, FOSL2)",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 14,
               label.size = 14,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_MES_CRC_FINAL.png"), 
#    height = 5900, width = 6470, res = 400)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "genes_of_interest1",
               plot.title = "MES CRC factors",
               plot.subtitle = "(YAP1, WWTR1, JUN, FOSL1, FOSL2)",
               order = TRUE,
               pt.size = 2,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 60,
               label.size = 55,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "",
               legend.position = "right",
               legend.length = 20,
               legend.width = 4)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_genes_of_interest_shuffle.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "genes_of_interest1",
               plot.title = "Gene module score",
               plot.subtitle = "YAP1, WWTR1, JUN, FOSL1 and FOSL2",
           #   order = FALSE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 36,
               label.size = 36,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()
```

```{r plot genes of interest with ordering by expression}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_genes_of_interest_wOrderingByExpression.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "genes_of_interest1",
               plot.title = "Gene module score",
               plot.subtitle = "YAP1, WWTR1, JUN, FOSL1 and FOSL2",
               order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 14,
               label.size = 14,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_moduleScore_genes_of_interest_wOrderingByExpression.png"), 
#    height = 500, width = 400)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "genes_of_interest1",
               plot.title = "Gene module score",
               plot.subtitle = "YAP1, WWTR1, JUN, FOSL1 and FOSL2",
               order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               font.size = 20,
               label.size = 20,
               min.cutoff = 0, 
               max.cutoff = 3,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()
```


## ___Core MES genes: Gene expression___

```{r fig.height=5, fig.width=4}
# plot expression of individual genes
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_MES_markers_COL4A1.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "COL4A1",
               plot.title = "core MES markers",
               plot.subtitle = "COL4A1 gene expression",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 3,
               font.size = 14,
               label.size = 14,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_MES_markers_COL4A1.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "COL4A1",
               plot.title = "core MES markers",
               plot.subtitle = "COL4A1 gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 3,
               font.size = 36,
               label.size = 36,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()


#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_MES_markers_FN1.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "FN1",
               plot.title = "core MES markers",
               plot.subtitle = "VIM gene expression",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 3,
               font.size = 14,
               label.size = 14,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_MES_markers_FN1.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "FN1",
               plot.title = "core MES markers",
               plot.subtitle = "VIM gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 3,
               font.size = 40,
               label.size = 40,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()


#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_MES_markers_VIM.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "VIM",
               plot.title = "core MES markers",
               plot.subtitle = "VIM gene expression",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 3,
               font.size = 14,
               label.size = 14,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_MES_markers_VIM.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "VIM",
               plot.title = "core MES markers",
               plot.subtitle = "VIM gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 3,
               font.size = 40,
               label.size = 40,
               legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_MES_markers_CD44.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "CD44",
               plot.title = "core MES markers",
               plot.subtitle = "CD44 gene expression",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 3,
               font.size = 14,
               label.size = 14,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_MES_markers_CD44.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "CD44",
               plot.title = "core MES markers",
               plot.subtitle = "CD44 gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 3,
               font.size = 40,
               label.size = 40,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()
```

## ___Core ADR genes: Gene expression___

```{r PHOX2B}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_ADR_markers_PHOX2B.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "PHOX2B",
               plot.title = "core ADR markers",
               plot.subtitle = "PHOX2B gene expression",
               order = TRUE,
               raster = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 4,
               font.size = 16,
               label.size = 16,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"NB_HR_patients_complete/figures/NB_HR_patients_complete_core_ADR_markers_PHOX2B.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "PHOX2B",
               plot.title = "core ADR markers",
               plot.subtitle = "PHOX2B gene expression",
               order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
              max.cutoff = 4,
               font.size = 40,
               label.size = 40,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()
```

```{r DBH}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_ADR_markers_DBH.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "DBH",
               plot.title = "core ADR markers",
               plot.subtitle = "DBH gene expression",
               raster = TRUE,
               order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 2,
               font.size = 16,
               label.size = 16,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_ADR_markers_DBH.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "DBH",
               plot.title = "core ADR markers",
               plot.subtitle = "DBH gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 2,
               font.size = 40,
               label.size = 40,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()
```

```{r GATA3}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_ADR_markers_GATA3.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "GATA3",
               plot.title = "core ADR markers",
               plot.subtitle = "GATA3 gene expression",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 2,
               font.size = 14,
               label.size = 14,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_ADR_markers_GATA3.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "GATA3",
               plot.title = "core ADR markers",
               plot.subtitle = "GATA3 gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 2,
               font.size = 40,
               label.size = 40,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()
```

```{r TH}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_ADR_markers_TH.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "TH",
               plot.title = "core ADR markers",
               plot.subtitle = "TH gene expression",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 1,
               font.size = 14,
               label.size = 14,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_ADR_markers_TH.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "TH",
               plot.title = "core ADR markers",
               plot.subtitle = "TH gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 1,
               font.size = 40,
               label.size = 40,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()
```

## ___Core SYM genes: Gene expression___

```{r MKI67}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_SYM_markers_MKI67.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "MKI67",
               plot.title = "core SYM markers",
               plot.subtitle = "MKI67 gene expression",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 1,
               font.size = 14,
               label.size = 14,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_SYM_markers_MKI67.png"), 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "MKI67",
               plot.title = "core SYM markers",
               plot.subtitle = "MKI67 gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 1,
               font.size = 40,
               label.size = 40,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()
```

```{r TOP2A}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_SYM_markers_TOP2A.pdf"), 
#    height = 5, width = 4)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "TOP2A",
               plot.title = "core SYM markers",
               plot.subtitle = "TOP2A gene expression",
               #order = TRUE,
               pt.size = 0.25,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 2,
               font.size = 14,
               label.size = 14,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 10)
#dev.off()

##png(file = paste0(params$out,"/NB_HR_patients_complete/figures/NB_HR_patients_complete_core_SYM_markers_TOP2A.png") 
#    height =2000, width =1600,res = 200)
do_FeaturePlot(sample = NB_HR_patients_complete, 
               features = "TOP2A",
               plot.title = "core SYM markers",
               plot.subtitle = "TOP2A gene expression",
               #order = TRUE,
               pt.size = 0.5,
               viridis.direction = -1,
               viridis.palette = "D",
               min.cutoff = 0, 
               max.cutoff = 2,
               font.size = 40,
               label.size = 40,
               #legend.title.face = "bold",
               axis.text.face = "bold",
               plot.title.face = "bold",
               plot.axes = TRUE,
               legend.title = "Gene expression",
               legend.length = 20)
#dev.off()
```


## ___ViolinPlot and DotPlot___

```{r eval=FALSE, include=FALSE}
for (name in names(gene.lists)) {
  print(paste0("Plotting module score for: ", name))
  for (i in 1:length(gene.lists)) {
    plot <- FeaturePlot(object = NB_HR_patients_complete, 
                        features = paste0(name,"1"), 
                        col =  c("gray","orange","red"),
                        order = TRUE)
    print(plot)
  }
}
```



```{r fig.height=8, fig.width=8}
module_scores_v0 <- c("Strohmenger2024 MES" ="soren2024_MES_genes_v21",
                    "Strohmenger2024 ADR" = "soren2024_ADR_genes_v21")

module_scores <- c("Strohmenger2024 MES" ="soren2024_MES_genes_v21",
                    "Van Groeningen MES" = "vangroningen2017_MES_genes1",
                    "Boeva NCC" =  "boeva2021_NCC_genes1",
                    "Strohmenger2024 ADR" = "soren2024_ADR_genes_v21",
                    "Van Groeningen ADR" = "vangroningen2017_ADR_genes1",
                    "Boeva ADR" = "boeva2021_ADR_genes1")

module_scores_v2 <- c("soren2024_MES_genes_v21","soren2024_ADR_genes_v21",
                  "vangroningen2017_MES_genes1","vangroningen2017_ADR_genes1",
                  "boeva2021_NCC_genes1","boeva2021_ADR_genes1",
                  "thirant2023_MES_genes1","thirant2023_ADR_SYM_genes1")


module_scores_v3 <- c("soren2024_MES_genes1","soren2024_ADR_genes1",
                  "vangroningen2017_MES_genes1","vangroningen2017_ADR_genes1",
                  "boeva2021_NCC_genes1","boeva2021_ADR_genes1")

module_scores_v4 <- c("soren2024_MES_genes_v31","soren2024_ADR_genes_v31",
                  "vangroningen2017_MES_genes1","vangroningen2017_ADR_genes1",
                  "boeva2021_NCC_genes1","boeva2021_ADR_genes1")


DotPlot(NB_HR_patients_complete, 
        features = module_scores,
        cols = c("lightblue","red")) +
        theme(axis.text.x = element_text(angle = 45,hjust=1, size = 16)) + 
        RotatedAxis() + scale_size(range = c(1,8)) 


DotPlot(NB_HR_patients_complete, 
        features = module_scores_v0,
        cols = c("lightblue","red")) +
        theme(axis.text.x = element_text(angle = 45,hjust=1, size = 16)) + 
        RotatedAxis() + scale_size(range = c(3,8)) 
```


```{r dotplot module scores}
Clustered_DotPlot(NB_HR_patients_complete, 
                  features = module_scores,
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v2,
                  exp_color_min = -1,
                  exp_color_middle = 0,
                  exp_color_max = 1,
                  cluster_ident = TRUE) 

Clustered_DotPlot(NB_HR_patients_complete, 
                  features = module_scores,
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v2,
                  exp_color_min = -1,
                  exp_color_middle = 0,
                  exp_color_max = 1,
                  row_label_size = 10,
                  cluster_ident = TRUE) 


Clustered_DotPlot(NB_HR_patients_complete, 
                  features = module_scores_v2,
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v2,
                  exp_color_min = -1,
                  exp_color_middle = 0,
                  exp_color_max = 1,
                  cluster_ident = TRUE) 


Clustered_DotPlot(NB_HR_patients_complete, 
                  features = module_scores,
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v3,
                  exp_color_min = -1,
                  exp_color_middle = 0,
                  exp_color_max = 1,
                  cluster_ident = TRUE) 

Clustered_DotPlot(NB_HR_patients_complete, 
                  features = module_scores_v3,
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v2,
                  exp_color_min = -1,
                  exp_color_middle = 0,
                  exp_color_max = 1,
                  cluster_ident = TRUE) 

Clustered_DotPlot(NB_HR_patients_complete, 
                  features = module_scores_v4,
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  show_parent_dend_line = TRUE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v2,
                  exp_color_min = -1,
                  exp_color_middle = 0,
                  exp_color_max = 1,
                  cluster_ident = TRUE) 

```

```{r violin plots}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/ViolinPlot_GenesOfInterest.pdf"), 
#    height = 15, width = 20)
VlnPlot(NB_HR_patients_complete, 
        features = dotplot_genes, 
        log = TRUE, pt.size = 0)
#dev.off()
```

## ___Dotplots___

```{r dotplot}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/DotPlot_GenesOfInterest_v12-2024.pdf"), 
#    height = 15, width = 20)
DotPlot(NB_HR_patients_complete, 
        features = dotplot_genes,
        cols = c("blue1","red")) + 
  theme(axis.text.x = element_text(angle = 90, hjust=1, size = 16))
  theme(axis.text.y = element_text(angle = 45, hjust=1))
#dev.off()
```


```{r dotplot}
#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/DotPlot_v1_FINAL.pdf"),
#    width = 5.72, height = 3.79)
Clustered_DotPlot(NB_HR_patients_complete, 
                  plot_km_elbow = FALSE,
                  features = dotplot_genes,
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v3,
                  exp_color_min = -2,
                  exp_color_middle = 0,
                  exp_color_max = 2,
                  row_label_size = 12,
                  column_label_size = 12,
                  legend_label_size = 12,
                  x_lab_rotate = 90,
                  legend_title_size = 12)
#dev.off()

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/DotPlot_v2_FINAL.pdf"),
#    width = 5.72, height = 3.79)
Clustered_DotPlot(NB_HR_patients_complete, 
                  plot_km_elbow = FALSE,
                  features = dotplot_genes_v2, # removed SCP markers
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v3,
                  exp_color_min = -2,
                  exp_color_middle = 0,
                  exp_color_max = 2,
                  row_label_size = 12,
                  column_label_size = 12,
                  legend_label_size = 12,
                  x_lab_rotate = 90,
                  legend_title_size = 12)
#dev.off()

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/DotPlot_v3_FINAL.pdf"),
#    width = 5.72, height = 3.79)
Clustered_DotPlot(NB_HR_patients_complete, 
                  plot_km_elbow = FALSE,
                  features = dotplot_genes_v3, # add one SCP marker
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v3,
                  exp_color_min = -2,
                  exp_color_middle = 0,
                  exp_color_max = 2,
                  row_label_size = 12,
                  column_label_size = 12,
                  legend_label_size = 12,
                  x_lab_rotate = 90,
                  legend_title_size = 12)
#dev.off()

#pdf(file = paste0(params$out,"/NB_HR_patients_complete/figures/DotPlot_v12-2024.pdf"),
#    width = 5.72, height = 3.79)
Clustered_DotPlot(NB_HR_patients_complete, 
                  plot_km_elbow = FALSE,
                  features = dotplot_genes_v4, # PLP1 instead of S100B
                  colors_use_exp = c("lightblue","red"),
                  raster = FALSE,
                  seed = 2805,
                  colors_use_idents = COLOR_CODE_v3,
                  exp_color_min = -2,
                  exp_color_middle = 0,
                  exp_color_max = 2,
                  row_label_size = 12,
                  column_label_size = 12,
                  legend_label_size = 12,
                  x_lab_rotate = 90,
                  legend_title_size = 12)
#dev.off()
```


# ___Differential abundance testing for tumor clusters___

```{r eval=FALSE, include=FALSE}
NB_HR_patients_complete <- readRDS(file = paste0(params$mnt_out, "/patel2024/Rds/patel2024_NB-HR_tumorCells.Rds"))
```

```{r eval=FALSE, include=FALSE}
NB_HR_patients_complete@meta.data$sample_name <- factor(NB_HR_patients_complete@meta.data$orig.ident)
NB_HR_patients_complete@meta.data$Treatment_Status <- factor(NB_HR_patients_complete@meta.data$Treatment_Status)

# add detailed treatment comparison groups

Idents(NB_HR_patients_complete) <- NB_HR_patients_complete@meta.data$treatment 

NB_HR_patients_complete <- RenameIdents(NB_HR_patients_complete, 
                     "untreated" = "at_diagnosis",
                     "treated, chemotherapy" = "after_chemo",
                     "treated, chemo-immunotherapy" = "after_chemo_immuno")

# "untreated"
# "treated, chemotherapy"
# "treated, chemo-immunotherapy"

NB_HR_patients_complete$treatment_groups_detailed <- NB_HR_patients_complete@active.ident

# add short treatment groups

Idents(NB_HR_patients_complete) <- NB_HR_patients_complete@meta.data$Treatment_Status 

NB_HR_patients_complete <- RenameIdents(NB_HR_patients_complete, 
                     "no treatment" = "untreated",
                     "sample collected after several courses of induction chemotherapy" = "treated")

NB_HR_patients_complete$treatment_groups <- NB_HR_patients_complete@active.ident


# add short MYCN status comparisons [so i can add MYCN as a covariate]

Idents(NB_HR_patients_complete) <- NB_HR_patients_complete@meta.data$MYCN

NB_HR_patients_complete <- RenameIdents(NB_HR_patients_complete, 
                     "amplified" = "",
                     "non-amplified" = "")

NB_HR_patients_complete$MYCN_groups <- NB_HR_patients_complete@active.ident

table(NB_HR_patients_complete$MYCN_groups, 
      NB_HR_patients_complete$treatment_groups)

saveRDS(NB_HR_patients_complete, 
        file =  paste0(params$nb_tumors,
                  "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR_tumor_cluster.Rds"))
```

```{r}
NB_HR_patients_complete <- readRDS(file = 
                               paste0(params$nb_tumors,
                                      "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR_tumor_cluster.Rds"))
```

```{r fig.height=8, fig.width=15}
Idents(NB_HR_patients_complete) <- NB_HR_patients_complete$liger_clusters

DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST, 
        raster = FALSE,
        shuffle = FALSE,
        label.size = 40, 
        split.by = "treatment_groups")

Idents(NB_HR_patients_complete) <- NB_HR_patients_complete$treatment_groups

DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        raster = FALSE,
        shuffle = FALSE,
        label.size = 40, 
        split.by = "INRGSS_Stage") 

# only one stage L and one stage Ms high-risk NB patient, both untreated
# i think we should remove them as this will only hinder the analysis
```

```{r}
filter_by_stage <- NB_HR_patients_complete$INRGSS_Stage == "M"
NB_HR_patients_complete_stage_M_only <- NB_HR_patients_complete[,filter_by_stage]

NB_HR_patients_complete_stage_M_only$sample_name <- droplevels(NB_HR_patients_complete_stage_M_only$sample_name)

NB_HR_patients_complete_stage_M_only$sample_name <- factor(NB_HR_patients_complete_stage_M_only$sample_name)
NB_HR_patients_complete_stage_M_only$MYCN <- factor(NB_HR_patients_complete_stage_M_only$MYCN)

```


## ___Untreated vs. treated HR-NB stage M primary tumors___

### ___BOX PLOTS___

```{r STATS test - boxplots}
Idents(NB_HR_patients_complete_stage_M_only) <- NB_HR_patients_complete_stage_M_only$liger_clusters

DimPlot(NB_HR_patients_complete_stage_M_only,
        group.by = "liger_clusters",
        split.by = "treatment_groups",
        cols = COLOR_CODE_CLUST_v2)

percent_pergroup <- table(NB_HR_patients_complete_stage_M_only$liger_clusters,
                          NB_HR_patients_complete_stage_M_only$treatment_groups)
percent_pergroup

percent_persample <- table(NB_HR_patients_complete_stage_M_only$liger_clusters,
                          NB_HR_patients_complete_stage_M_only$sample_name)

barplot(percent_pergroup,
        col = COLOR_CODE_CLUST_v2, 
        border = "white", 
        xlab = "liger_clusters")

# Transform this data in %
data_percentage <- apply(percent_pergroup, 2, 
                         function(x){x*100/sum(x,na.rm=T)})
 

data_percentage_sample <- apply(percent_persample, 2, 
                         function(x){x*100/sum(x,na.rm=T)})

# Make a stacked barplot in %
barplot(data_percentage_sample, 
        col = COLOR_CODE_CLUST_v2, 
        border = "white", 
        xlab = "liger_clusters")

table_per_group <- as.data.frame(percent_pergroup)
table_per_group$group <- table_per_group$Var2
table_per_group

######################################################################

library(Seurat)
library(dplyr)
library(tidyr)

# Aggregate cell counts by sample and cluster
cluster_counts <- table(NB_HR_patients_complete_stage_M_only$liger_clusters, 
                        NB_HR_patients_complete_stage_M_only$sample_name)


# Calculate total cell counts per sample
total_lib_size <- colSums(cluster_counts[,-1])  # Exclude header
total_lib_size <- as.data.frame(total_lib_size)
total_lib_size$sample_name <- rownames(total_lib_size) 

# Convert to %
proportions <- sweep(cluster_counts, 2, colSums(cluster_counts), FUN = "/")

# Add metadata for conditions
treatment_groups <- unique(data.frame("treatment_groups" = 
  NB_HR_patients_complete_stage_M_only$treatment_groups, "sample_name" =
  NB_HR_patients_complete_stage_M_only$sample_name, "MYCN" =
  NB_HR_patients_complete_stage_M_only$MYCN))
 
treatment_groups_wCellCounts <- merge(treatment_groups,
      total_lib_size,
      by ="sample_name")

# Convert to long format
library(reshape2)
cell_counts_long <- melt(cluster_counts)
colnames(cell_counts_long) <- c("liger_clusters","sample_name","percent_cells")

# Merge with metadata
counts_merged <- cell_counts_long %>%
  left_join(treatment_groups_wCellCounts, by="sample_name")

# View the combined data
print(counts_merged)

#####################################################################
# GLM glm_fit to account for decreasing tumor count after treatment #
#####################################################################

counts_merged$treatment_groups  <- factor(counts_merged$treatment_groups, 
                                          levels = c("untreated", "treated"))

clust_4 <- filter(counts_merged, liger_clusters == "4")

# Fit GLM
glm_fit <- glm(
  percent_cells / total_lib_size ~ 0 + treatment_groups * MYCN,
  family = binomial(link = "logit"), 
  weights = total_lib_size,
  data = clust_4
)

# Summarize the glm_fit
summary(glm_fit)


# Predicted proportions for each condition
predicted_data <- data.frame(treatment_groups = clust_4$treatment_groups, MYCN = clust_4$MYCN)
predicted <- predict(glm_fit, newdata = predicted_data, type = "response")
predicted_data$Predicted_Proportion <- predicted

print(predicted_data)


library(ggplot2)

# Add observed proportions to the data
predicted_data$Observed_Proportion <- clust_4$percent_cells / clust_4$total_lib_size

# Plot
ggplot(predicted_data, aes(x = treatment_groups, y = Observed_Proportion)) +
  geom_point(size = 3) +
  geom_line(data = predicted_data, aes(x = treatment_groups, y = Predicted_Proportion), 
            group = 1, color = "blue") +
  labs(title = "Change in Cell Type Abundance", y = "Proportion of cluster 4")

ggplot(predicted_data, aes(x = MYCN, y = Observed_Proportion)) +
  geom_point(size = 3) +
  geom_line(data = predicted_data, aes(x = MYCN, y = Predicted_Proportion), 
            group = 1, color = "blue") +
  labs(title = "Change in Cell Type Abundance", y = "Proportion of cluster 4")

# GLM Goodness-of-Fit: Deviance and Pearson Chi-Square Test
# Deviance test:
deviance_test <- glm_fit$deviance
deviance_df <- glm_fit$df.residual
p_value_deviance <- 1 - pchisq(deviance_test, deviance_df)
cat("Deviance test p-value:", p_value_deviance, "\n")

# Pearson Chi-Square test
pearson_chi2 <- sum((glm_fit$y - fitted(glm_fit))^2 / fitted(glm_fit))
pearson_df <- glm_fit$df.residual
p_value_pearson <- 1 - pchisq(pearson_chi2, pearson_df)
cat("Pearson Chi-Square test p-value:", p_value_pearson, "\n")

# 3. Residual Analysis
# Plot residuals vs fitted values
plot(fitted(glm_fit), residuals(glm_fit), 
     main = "Residuals vs Fitted", 
     xlab = "Fitted values", 
     ylab = "Residuals")
abline(h = 0, col = "red")

# QQ-plot of residuals
qqnorm(residuals(glm_fit))
qqline(residuals(glm_fit), col = "red")

# Check for overdispersion (optional)
overdispersion <- sum(residuals(glm_fit, type = "pearson")^2) / glm_fit$df.residual
cat("Overdispersion statistic:", overdispersion, "\n")

####################################
# Beta-binomial model using vglm() #
####################################

group <- ifelse(clust_4$treatment_groups == "treated", 1, 0)
clust_4$group <- group

clust_4$group  <- factor(clust_4$group,levels = c("1", "0"))


# Fit the beta-binomial model using 'betareg'
# Here, we model the abundance of a specific cell type as a function of the treatment status
library("betareg")
model_beta_binomial <- betareg(percent_cells ~ treatment_groups + MYCN | 1, data = clust_4)
summary(model_beta_binomial)

# Extract coefficients
coef(model_beta_binomial)

# Confidence intervals for coefficients
confint(model_beta_binomial)

# Predicted probabilities
predicted_data <- data.frame(treatment_groups = clust_4$treatment_groups, MYCN = clust_4$MYCN)
predicted <- predict(model_beta_binomial, type = "response")
predicted_data <- predicted_data[-1,]
predicted_data$Predicted_Proportion <- predicted

print(predicted_data)


# Add observed proportions to the data
clust_4 <- clust_4[-1,]
predicted_data$Observed_Proportion <- clust_4$percent_cells 

# Plot
ggplot(predicted_data, aes(x = treatment_groups, y = Observed_Proportion)) +
  geom_point(size = 3) +
  geom_line(data = predicted_data, aes(x = treatment_groups, y = Predicted_Proportion), 
            group = 1, color = "blue") +
  labs(title = "Change in Cell Type Abundance", y = "Proportion of cluster 4")

ggplot(predicted_data, aes(x = MYCN, y = Observed_Proportion)) +
  geom_point(size = 3) +
  geom_line(data = predicted_data, aes(x = MYCN, y = Predicted_Proportion), 
            group = 1, color = "blue") +
  labs(title = "Change in Cell Type Abundance", y = "Proportion of cluster 4")

# Test for DA using a statistical test (e.g., Wilcoxon rank-sum test)

# plot boxplot

# CLUSTER 4
clust_4 <- filter(counts_merged, liger_clusters == "4")
clust_4$comparison_groups <- paste0(clust_4$treatment_groups,"_",clust_4$MYCN)


# perform wilcox testing
# The Wilcoxon rank sum test is a non-parametric alternative to the independent two samples t-test for comparing two independent groups of samples, in the situation where the data are not normally distributed. Synonymous: Mann-Whitney test
# assumes the shape of the distribution is symmetrical!!!!!!

filt1 <- clust_4$group == 1
groupA <- clust_4[filt1,]

filt2 <- clust_4$group == 0
groupB <- clust_4[filt2,]

wilcox.test(x = groupA$percent_cells, 
            y = groupB$percent_cells,
            paired = FALSE) #  Wilcoxon rank sum test (Mann-Whitney test)

##############################################################################

# clust_4_and_6
filt1 <- clust_4_and_6$treatment_groups == "untreated"
groupA <- clust_4_and_6[filt1,]

filt2 <- clust_4_and_6$treatment_groups == "treated"
groupB <- clust_4_and_6[filt2,]

wilcox.test(x = groupA$percent_cells, 
            y = groupB$percent_cells,
            paired = FALSE) #  Wilcoxon rank sum test (Mann-Whitney test)

# separate statistical test for nMNA and MNNA samples
# untreated
filt1 <- groupA$MYCN == "non-amplified"
groupA_nMNA <- groupA[filt1,]
filt2 <- groupA$MYCN == "amplified"
groupA_MNA <- groupA[filt2,]

# treated
filt1 <- groupB$MYCN == "non-amplified"
groupB_nMNA <- groupB[filt1,]
filt2 <- groupB$MYCN == "amplified"
groupB_MNA <- groupB[filt2,]

# nMNA treated vs. untreated
wilcox.test(x = groupA_nMNA$percent_cells, 
            y = groupB_nMNA$percent_cells,
            paired = FALSE) #  Wilcoxon rank sum test (Mann-Whitney test)

# MNA treated vs. untreated
wilcox.test(x = groupA_MNA$percent_cells, 
            y = groupB_MNA$percent_cells,
            paired = FALSE) #  Wilcoxon rank sum test (Mann-Whitney test)

# ALTERNATIVE STRATEGY: Fisher's exact test for small sample sizes and rare outcomes
# i think this one is more appropriate for this particular data

fisher.test(table(clust_4_and_6$percent_cells, clust_4_and_6$treatment_groups))

# nMNA
filt1 <- clust_4_and_6$MYCN == "non-amplified"
nMNA <- clust_4_and_6[filt1,]

fisher.test(table(nMNA$percent_cells, nMNA$treatment_groups))

# MNA
filt1 <- clust_4_and_6$MYCN == "amplified"
MNA <- clust_4_and_6[filt1,]

fisher.test(table(MNA$percent_cells, MNA$treatment_groups))

###############################################################################

# clust_4_and_6_and_11
filt1 <- clust_4_and_6_and_11$treatment_groups == "untreated"
groupA <- clust_4_and_6_and_11[filt1,]

filt2 <- clust_4_and_6_and_11$treatment_groups == "treated"
groupB <- clust_4_and_6_and_11[filt2,]

wilcox.test(x = groupA$percent_cells, 
            y = groupB$percent_cells,
            paired = FALSE) #  Wilcoxon rank sum test (Mann-Whitney test)



library(ggsignif)
boxplot <- clust_4 %>%
  ggplot( aes(x = treatment_groups , y = percent_cells, fill = treatment_groups)) +
    geom_violin() +  geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 4") +
    xlab("") + ylab("Proportion of cells") +
  geom_signif(comparisons = list(c("untreated", "treated"),
              map_signif_level = FALSE))

pdf(file = paste0(params$out,"patel2024/figures/","cluster4_boxplot.pdf"))
boxplot
dev.off()

# CLUSTER 6
clust_6 <- filter(counts_merged, liger_clusters == "6")


# perform wilcox testing
p_value1<- wilcox.test(percent_cells ~ treatment_groups, 
                        data = subset(clust_6, 
                                      treatment_groups %in% c("untreated", "treated")))$p.value

library(ggsignif)
boxplot <- clust_6 %>%
  ggplot( aes(x = treatment_groups , y = percent_cells, fill = treatment_groups)) +
    geom_violin() +  geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 6") +
    xlab("") + ylab("Proportion of cells") +
  geom_signif(comparisons = list(c("untreated", "treated"),
              map_signif_level = FALSE))

pdf(file = paste0(params$out,"patel2024/figures/","cluster6_boxplot.pdf"))
boxplot
dev.off()

# CLUSTER 11
clust_11 <- filter(counts_merged, liger_clusters == "11")

# perform wilcox testing
p_value1<- wilcox.test(percent_cells ~ treatment_groups, 
                        data = subset(clust_11, 
                                      treatment_groups %in% c("untreated", "treated")))$p.value


library(ggsignif)
boxplot <- clust_11 %>%
  ggplot( aes(x = treatment_groups , y = percent_cells, fill = treatment_groups)) +
    geom_violin() +  geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.11) +
    theme(legend.position="none",
          axis.text = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 11") +
    xlab("") + ylab("Proportion of cells") +
  geom_signif(comparisons = list(c("untreated", "treated"),
              map_signif_level = FALSE))

pdf(file = paste0(params$out,"patel2024/figures/","cluster11_boxplot.pdf"))
boxplot
dev.off()


###################################### MYCN #####################################

# Aggregate cell counts by sample and cluster
cluster_counts <- table(NB_HR_patients_complete_stage_M_only$liger_clusters,
                        NB_HR_patients_complete_stage_M_only$sample_name)

# Convert to %
proportions <- sweep(cluster_counts, 2, colSums(cluster_counts), FUN = "/")

# Add metadata for conditions
treatment_groups <- unique(data.frame(
  "MYCN" = NB_HR_patients_complete_stage_M_only$MYCN, 
  "sample_name" = NB_HR_patients_complete_stage_M_only$sample_name,
  "treatment_groups" =  NB_HR_patients_complete_stage_M_only$treatment_groups))


# Convert to long format
cell_counts_long <- melt(proportions)
colnames(cell_counts_long) <- c("liger_clusters","sample_name","percent_cells")

# Merge with metadata
counts_merged <- cell_counts_long %>%
  left_join(treatment_groups, by="sample_name")

# View the combined data
print(counts_merged)


# Test for DA using a statistical test (e.g., Wilcoxon rank-sum test)

# plot boxplot

# CLUSTER 4
clust_4 <- filter(counts_merged, liger_clusters == "4")

library(ggsignif)
library(viridis)

boxplot <- clust_4 %>%
  ggplot( aes(x = MYCN , y = percent_cells, fill = MYCN)) +
    geom_violin() + geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 4") +
    xlab("") + ylab("Proportion of cells") +
  geom_signif(comparisons = list(c("amplified", "non-amplified"),
              map_signif_level = FALSE))

pdf(file = paste0(params$out,"patel2024/figures/","cluster4_boxplot_MYCN.pdf"))
boxplot
dev.off()


pdf(file = paste0(params$out,"patel2024/figures/","cluster4_boxplot_MYCNsubgroups.pdf"))
clust_4 %>%
    ggplot( aes(x = comparison_groups  , y = percent_cells, fill = comparison_groups )) +
    geom_violin() +  geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90,size = 18),
          axis.text.y = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 4") +
    xlab("") + ylab("Proportion of cells") 
dev.off()

#treatment_groups
clust_4_untreated <- filter(counts_merged, treatment_groups == "untreated")
clust_4_treated <- filter(counts_merged, treatment_groups == "treated")

library(ggsignif)
boxplot <- clust_4_untreated %>%
  ggplot( aes(x = MYCN , y = percent_cells, fill = MYCN)) +
    geom_violin() + geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 4") +
    xlab("") + ylab("Proportion of cells") +
  geom_signif(comparisons = list(c("amplified", "non-amplified"),
              map_signif_level = FALSE))

clust_4_treated %>%
  ggplot( aes(x = MYCN , y = percent_cells, fill = MYCN)) +
    geom_violin() + geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 4") +
    xlab("") + ylab("Proportion of cells") +
  geom_signif(comparisons = list(c("amplified", "non-amplified"),
              map_signif_level = FALSE))

pdf(file = paste0(params$out,"patel2024/figures/","cluster4_boxplot_MYCN.pdf"))
boxplot
dev.off()

# CLUSTER 4 + 6
proportions <- sweep(cluster_counts, 2, colSums(cluster_counts), FUN = "/")
proportions <- melt(proportions)
# Combine clusters 4 and 6

colnames(proportions) <- c("liger_clusters","sample_name","percent_cells")

combined_proportions_4and6 <- proportions %>%
  mutate(liger_clusters =
         ifelse(liger_clusters %in% c("4", "6"), "4_and_6",
                no = liger_clusters))

# Merge with metadata
counts_merged <- combined_proportions_4and6 %>%
  left_join(treatment_groups, by="sample_name")

clust_4_and_6 <- filter(counts_merged, liger_clusters == "4_and_6")

clust_4_and_6$comparison_groups <- paste0(clust_4_and_6$treatment_groups,"_",clust_4_and_6$MYCN)

clust_4_and_6_plot <- clust_4_and_6 %>%
  ggplot( aes(x = treatment_groups , y = percent_cells, fill = treatment_groups)) +
    geom_violin() + geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Clust 4+6") +
    xlab("") + ylab("Proportion of cells") +
  geom_signif(comparisons = list(c("untreated", "treated"),
              map_signif_level = FALSE))

pdf(file = paste0(params$out,"patel2024/figures/","cluster4_and_6_vlnPlot.pdf"))
clust_4_and_6_plot
dev.off()


# perform wilcox testing
wilcox.test(percent_cells ~ treatment_groups, 
            data = subset(clust_4_and_6, 
                          treatment_groups %in% c("untreated", "treated")))$p.value


pdf(file = paste0(params$out,"patel2024/figures/","cluster4_and_6_boxplot_MYCNsubgroups.pdf"))
clust_4_and_6 %>%
    ggplot( aes(x = comparison_groups  , y = percent_cells, fill = comparison_groups )) +
    geom_violin() +  geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90,size = 18),
          axis.text.y = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 4+6") +
    xlab("") + ylab("Proportion of cells") 
dev.off()

# perform wilcox testing
wilcox.test(percent_cells ~ comparison_groups, 
            data = subset(clust_4_and_6, 
                          comparison_groups %in% c("treated_amplified", "untreated_amplified")))$p.value

# perform wilcox testing
wilcox.test(percent_cells ~ comparison_groups, 
            data = subset(clust_4_and_6, 
                          comparison_groups %in% c("treated_non-amplified", "untreated_non-amplified")))$p.value

# CLUSTER 4 + 6 + 11

combined_proportions_4and6and11 <- proportions %>%
  mutate(liger_clusters =
         ifelse(liger_clusters %in% c("4", "6","11"), "4_and_6_and_11",
                no = liger_clusters))

# Merge with metadata
counts_merged <- combined_proportions_4and6and11 %>%
  left_join(treatment_groups, by="sample_name")

clust_4_and_6_and_11 <- filter(counts_merged, liger_clusters == "4_and_6_and_11")

clust_4_and_6_and_11$comparison_groups <- paste0(clust_4_and_6_and_11$treatment_groups,"_",clust_4_and_6_and_11$MYCN)

clust_4_and_6_and_11_plot <- clust_4_and_6_and_11 %>%
  ggplot( aes(x = treatment_groups , y = percent_cells, fill = treatment_groups)) +
    geom_violin() + geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Clust 4+6") +
    xlab("") + ylab("Proportion of cells") +
  geom_signif(comparisons = list(c("untreated", "treated"),
              map_signif_level = FALSE))

pdf(file = paste0(params$out,"patel2024/figures/","cluster4_and_6_and_11_vlnPlot.pdf"))
clust_4_and_6_and_11_plot
dev.off()


# perform wilcox testing
wilcox.test(percent_cells ~ treatment_groups, 
            data = subset(clust_4_and_6_and_11, 
                          treatment_groups %in% c("untreated", "treated")))$p.value


pdf(file = paste0(params$out,"patel2024/figures/","cluster_4_and_6_and_11_boxplot_MYCNsubgroups.pdf"))
clust_4_and_6_and_11 %>%
    ggplot( aes(x = comparison_groups  , y = percent_cells, fill = comparison_groups )) +
    geom_violin() +  geom_boxplot(width=0.1) +
    scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90,size = 18),
          axis.text.y = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          plot.title = element_text(size = 18)) +
    ggtitle("NB tumor clusters: Cluster 4+6+11") +
    xlab("") + ylab("Proportion of cells") 
dev.off()

# perform wilcox testing
wilcox.test(percent_cells ~ comparison_groups, 
            data = subset(clust_4_and_6, 
                          comparison_groups %in% c("treated_non-amplified", "untreated_non-amplified")))$p.value
```


```{r miloR - prepare data}
# MiloR doesn't use the "original" clusters but rather creates neighborhoods
# based on similar logFC (note the fold change here is the number of cells !!
# per cell type)

# 1. pre-processing
Idents(NB_HR_patients_complete_stage_M_only) <- NB_HR_patients_complete_stage_M_only$liger_clusters

NB_HR_patients_complete_stage_M_only <- RunPCA(NB_HR_patients_complete_stage_M_only,  
                            npcs = 50,
                            reduction.name = "PCA")
#NB_HR_patients_complete_stage_M_only <- RunUMAP(NB_HR_patients_complete_stage_M_only, 
#                             dims = 1:50,
#                             reduction = "PCA",
#                             reduction.name = "new_UMAP")

# 2. save as single cell experiment (SCE)
rna_milo <- Milo(as.SingleCellExperiment(NB_HR_patients_complete_stage_M_only))
rna_milo

# 3. add log counts
#logcounts(rna_milo) <- log(counts(rna_milo) + 1)

# 4. add PCA and UMAP reductions to the SCE 
reducedDim(rna_milo, "PCA") <- reducedDim(as.SingleCellExperiment(NB_HR_patients_complete_stage_M_only), 
                                           "PCA")
reducedDim(rna_milo, "UMAP") <- reducedDim(as.SingleCellExperiment(NB_HR_patients_complete_stage_M_only), 
                                           "UMAP")
# 5. QC / sanity check
plotUMAP(rna_milo, 
         colour_by="treatment_groups") + 
  plotUMAP(rna_milo, 
           colour_by="sample_name")

# 6. Constructing kNN graph 
rna_milo <- buildGraph(rna_milo, 
                       k = 10, 
                       d = 30)

# MiloR tool defined a neighbourhood of an index cell as the group of cells 
# connected by an edge in the KNN graph to the index cell. For efficiency, 
# testing for DA in the neighbourhood is not perfomed for every cell, but on 
# a subset of representative cells, using a KNN sampling algorithm used by 
# Gut et al. 2015.

# 7. Define cell neighbourhoods, with k = 10
rna_milo <- makeNhoods(rna_milo, 
                       prop = 0.1, 
                       k = 10, 
                       d = 30, 
                       refined = TRUE
                       ) 
#### refinement_scheme = "graph" -> Make nhoods using graph-only FASTER

plotNhoodSizeHist(rna_milo)

# 8. Counting cells per neighborhood
rna_milo <- countCells(rna_milo, 
                       meta.data = data.frame(colData(rna_milo)),
                       samples="sample_name") # make an "nhood x patients" counts matrix

# 9. Defining experimental design
rna_design <- data.frame(colData(rna_milo))[,c("sample_name", 
                                               "treatment_groups",
                                               "MYCN")]
rna_design <- distinct(rna_design)
rownames(rna_design) <- rna_design$sample_name
rna_design

# 10. Computing neighborhood connectivity [used to compute p-values]
rna_milo <- calcNhoodDistance(rna_milo, 
                              d = 30)
```

```{r miloR - DA testing}
# DA testing
da_results <- testNhoods(rna_milo, 
                         design = ~ 0 + treatment_groups + MYCN, # treatment * MYCN produced no hoods
                         design.df = rna_design)
head(da_results)

da_results %>%
  arrange(- SpatialFDR) %>%
  head() 

rna_milo <- buildNhoodGraph(rna_milo)

# QUALITY CHECK OF THE DIFFERENTIAL ABUNDANCE TESTING

# a. Neighborhoods plot
nh_graph_pl <- plotNhoodGraphDA(rna_milo, 
                                da_results, 
                                layout = "UMAP", 
                                alpha = 0.05) 

# b. Single-cell UMAP
umap_pl <- plotReducedDim(rna_milo, 
                          dimred = "UMAP", 
                          colour_by = "sample_name",
                          text_by = "treatment_groups", 
                          text_size = 3, 
                          point_size = 0.5) +
  guides(fill = "none")

# c. combined plots
umap_pl + nh_graph_pl +
  plot_layout(guides = "collect")

# d.  logFC vulcano plot
ggplot(da_results, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)
```

```{r miloR - results}
table(da_results$SpatialFDR < 0.1)

# annotate hood with original cell type annotation
da_results <- annotateNhoods(rna_milo, 
                             da_results, 
                             coldata_col = "liger_clusters")

# define cut-off for homogeneous vs. heterogeneous neighbourhoods
ggplot(da_results, aes(liger_clusters_fraction)) + geom_histogram(bins=500)

# remove hoods that contain mixed cell type populations (homogeneity above 80%)
da_results$cluster_fraction <- 
  ifelse(da_results$liger_clusters_fraction < 0.9, 
         "Mixed", da_results$liger_clusters)

# Automatic grouping of neighborhoods
da_results <- groupNhoods(rna_milo, 
                          da_results, 
                          da.fdr = 0.1,
                          max.lfc.delta = 10)
head(da_results)

plotNhoodGroups(rna_milo, da_results, layout = "UMAP")

# filter only significant populations (less than 1% false discovery rate and 
# fold change bigger than 5)
results <- da_results %>% 
  arrange(FDR) %>% filter(FDR < 0.1 & abs(logFC) >= 2) # based on the histogram

plotDAbeeswarm(da_results, group.by = "liger_clusters")
plotDAbeeswarm(da_results, group.by = "NhoodGroup")
```

```{r save results}
write.table(results, 
            file = 
            paste0(params$out, "patel2024/tables/DA_testing_treatedVSuntreated.csv"),
                   sep = "\t", quote = FALSE)

saveRDS(rna_milo, 
        file = 
          paste0(params$nb_tumors,
                      "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR_tumorCells_DAtesting.Rds"))
```

## ___Untreated vs. chemotherapy vs. chemo-immunotherapy___

Since I are comparing >2 groups, I need to use contrasts here!

```{r}
# miloR with contrasts
gc()

# 10. Defining experimental design
rna_design <- data.frame(colData(rna_milo))[,c("sample_name", 
                                               "treatment_groups",
                                               "anti_GD2",
                                               "Gender",
                                               "Age",
                                               "MYCN")]
rna_design <- distinct(rna_design)
rownames(rna_design) <- rna_design$sample_name
rna_design

# 11. Computing neighborhood connectivity [used to compute p-values]
rna_milo <- calcNhoodDistance(rna_milo, 
                              d = 30)
```

```{r untreated vs. (chemo + chemo/immuno)}
glm_fit <- glm_fit.matrix(~ 0 + treatment_groups_detailed + MYCN, data = rna_design)
ave.contrast <- c("treatment_groups_detailedat_diagnosis - 
                  ((treatment_groups_detailedafter_chemo + treatment_groups_detailedafter_chemo_immuno) / 2)")

mod.constrast <- makeContrasts(contrasts = ave.contrast, levels = glm_fit)
mod.constrast

da_results_ChemoVsImmuno <- testNhoods(rna_milo, 
                                                design = ~ 0 + treatment_groups_detailed + MYCN,
                                                design.df = rna_design, 
                                                glm_fit.contrasts = ave.contrast, 
                                                fdr.weighting = "graph-overlap")

table(da_results_ChemoVsImmuno$SpatialFDR < 0.01)

# annotate hood with original cluster annotations
da_results_ChemoVsImmuno <- annotateNhoods(rna_milo, 
                                          da_results_ChemoVsImmuno, 
                                          coldata_col = "liger_clusters")

# remove hoods that contained mixed cell type populations
da_results_ChemoVsImmuno$liger_clusters <- 
  ifelse(da_results_ChemoVsImmuno$liger_clusters_fraction < 0.9, 
         "Mixed", da_results_ChemoVsImmuno$liger_clusters)

# filter only significant populations (less than 1% false discovery rate and 
# fold change bigger than 5)
results <- da_results_ChemoVsImmuno %>% 
  arrange(FDR) %>% filter(FDR < 0.01 & abs(logFC) >= 5)
results

plotDAbeeswarm(results, 
               group.by = "liger_clusters")
```

```{r}
write.table(results, file = 
             paste0(params$out, "patel2024/tables/DA_testing_treatedVSuntreated_wContrasts.csv"),
                   sep = "\t", quote = FALSE)

saveRDS(rna_milo, 
        file = 
          paste0(params$nb_tumors,
                      "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR_tumorCells_DAtesting_wContrasts.Rds"))
```

```{r chemo vs. chemo/immuno}
glm_fit <- glm_fit.matrix(~ 0 + treatment_groups_detailed, data = rna_design)
ave.contrast <- c("treatment_groups_detailedafter_chemo - treatment_groups_detailedafter_chemo_immuno")

mod.constrast <- makeContrasts(contrasts = ave.contrast, levels = glm_fit)
mod.constrast

da_results_ChemoVsImmuno <- testNhoods(rna_milo, 
                                                design = ~ 0 + treatment_groups_detailed + Gender,
                                                design.df = rna_design, 
                                                glm_fit.contrasts = ave.contrast, 
                                                fdr.weighting = "graph-overlap")

table(da_results_ChemoVsImmuno$SpatialFDR < 0.1)

# annotate hood with original cluster annotations
da_results_ChemoVsImmuno <- annotateNhoods(rna_milo, 
                                          da_results_ChemoVsImmuno, 
                                          coldata_col = "liger_clusters")

# remove hoods that contained mixed cell type populations

# but first set cut-off
plotNhoodSizeHist(rna_milo)

da_results_ChemoVsImmuno$liger_clusters <- 
  ifelse(da_results_ChemoVsImmuno$liger_clusters_fraction < 0.5, 
         "Mixed", da_results_ChemoVsImmuno$liger_clusters)

# filter only significant populations (less than 1% false discovery rate and 
# fold change bigger than 5)
results <- da_results_ChemoVsImmuno %>% 
  arrange(FDR) %>% filter(FDR < 0.1 & abs(logFC) >= 1)

results <- results %>% arrange(-liger_clusters_fraction)
results

plotDAbeeswarm(results, 
               group.by = "liger_clusters")
```

```{r}
write.table(results, file = 
             paste0(params$out, "patel2024/tables/DA_testing_chemo-vs-immuno_wContrasts.csv"),
                   sep = "\t", quote = FALSE)

saveRDS(rna_milo, 
        file = 
          paste0(params$nb_tumors,
                      "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR_tumorCells_DAtesting_chemo-vs-immuno.Rds"))
```


# ___nhood markers___

Find hood markers, than aggregating the neighbourhood expression profiles by experimental samples (the same used for DA testing), by setting aggregate.samples=TRUE. This way single-cells will not be considered as “replicates” during DGE testing, and dispersion will be estimated between true biological replicates. 

```{r}

da_results_ChemoVsImmuno <- groupNhoods(rna_milo, 
                          da_results_ChemoVsImmuno,
                          da.fdr = 0.01,
                          max.lfc.delta = 10)

head(da_results_ChemoVsImmuno)
plotNhoodGroups(rna_milo, 
                da_results_ChemoVsImmuno, 
                layout="UMAP") 

da_nhood_markers <- findNhoodGroupMarkers(rna_milo, 
                                          da_results_ChemoVsImmuno, 
                                          subset.row = rownames(rna_milo)[1:10],
                                          aggregate.samples = TRUE, 
                                          sample_col = "sample_name")

rna_milo <- buildNhoodGraph(rna_milo)

plotDAbeeswarm(da_results, "NhoodGroup")

# CLUSTER 4 DEGs between treated and untreated tumor samples

ggplot(da_nhood_markers, 
       aes(logFC_4, -log10(adj.P.Val_5 ))) + 
  geom_point(alpha=0.5, size=0.5) +
  geom_hline(yintercept = 3)

set.seed(42)
markers <- da_nhood_markers$GeneID[da_nhood_markers$adj.P.Val_4 < 0.1 & 
                                     da_nhood_markers$logFC_4 > 0]

plotNhoodExpressionGroups(rna_milo, 
                          da_results_ChemoVsImmuno, 
                          features=intersect(rownames(rna_milo), markers[1:10]),
                          subset.nhoods = da_results_ChemoVsImmuno$NhoodGroup %in% '4', 
                          scale=TRUE,
                          grid.space = "fixed")

# DGE testing within a group
DGE <- testDiffExp(rna_milo, 
                          #da_results_ChemoVsImmuno, 
                          design = ~ treatment_groups, 
                          meta.data = data.frame(colData(rna_milo)),
                          subset.row = rownames(rna_milo))
                          #subset.nhoods=da_results$NhoodGroup=="4")

signif_genes <- DGE %>% arrange(FDR) %>% filter(FDR < 0.01 & abs(logFC) >= 1)
signif_genes
```


# ___Plots___

```{r}
DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST, 
        raster = TRUE,
        shuffle = TRUE,
        label.size = 16)

##pdf(file = paste0(params$out,"NB_HR_patients_complete/figures/NB_HR_patients_complete_UMAP_tumorCells.pdf"), 
##    height = 7, width = 8)
DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST, 
        raster = FALSE,
        shuffle = TRUE,
        label.size = 36)
#dev.off()

#pdf(file = paste0(params$out,"NB_HR_patients_complete/figures/NB_HR_patients_complete_UMAP_tumorCells_raster_TRUE.pdf"), 
#    height = 7, width = 8)
DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST, 
        raster = TRUE,
        shuffle = TRUE,
        label.size = 36)
#dev.off()

#pdf(file = paste0(params$out,"NB_HR_patients_complete/figures/NB_HR_patients_complete_UMAP_tumorCells_raster_TRUE_v2.pdf"), 
#    height = 7, width = 8)
DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST_v2, 
        raster = TRUE,
        shuffle = TRUE,
        label.size = 36)
#dev.off()


##png(file = paste0(params$out,"NB_HR_patients_complete/figures/NB_HR_patients_complete_UMAP_tumorCells_v2.png"),
#    height = 700, width = 800,res = 200)
DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST_v2, 
        raster = FALSE,
        shuffle = TRUE,
        label.size = 40)
#dev.off()


##png(file = paste0(params$out,"NB_HR_patients_complete/figures/NB_HR_patients_complete_UMAP_tumorCells.png"),
#    height = 700, width = 800,res = 200)
DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST, 
        raster = FALSE,
        shuffle = TRUE,
        label.size = 40)
#dev.off()
```


# ___Alternative DA testing method___

```{r}
library(DAseq)
python2use <- "/home/rstudio/.virtualenvs/r-reticulate/bin/python"

# 20 patients with NB histology, high-risk tumors and information on treatment
NB_HR_patients_complete <- readRDS(file = paste0(params$out,
                                  "patel2024/Rds/patel2024_NB-HR_tumor_cluster.Rds"))

filter_by_treatment <- NB_HR_patients_complete$Treatment_Status  == "no treatment"
NB_HR_patients_complete <- NB_HR_patients_complete[,filter_by_treatment]
table(NB_HR_patients_complete$Treatment_Status)


# Create a new column based on the existing column
NB_HR_patients_complete$treatment_groups <- ifelse(NB_HR_patients_complete$Treatment_Status == "no treatment", "untreated", "treated")

# Verify the new column has been added
table(NB_HR_patients_complete$treatment_groups)

labels_treat <- X.label.info[X.label.info$condition == "treated", "treatment_groups"]
labels_untrea <- X.label.info[X.label.info$condition == "untreated", "treatment_groups"]

da_cells <- getDAcells(  
    X = NB_HR_patients_complete,
    cell.labels = NB_HR_patients_complete$treatment_groups,
    labels.1 = labels_treat,
    labels.2 = labels_untrea,
    k.vector = seq(50, 500, 50))
```


# ___SessionInfo___

```{r}
sessionInfo()
```

