---
title: "MES project: Figures for paper revision"
author: "Sara Wernig-Zorc"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
    collapsed: no
    code_folding: hide
    highlight: pygments
    theme: spacelab
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '5'
params:
  input: "/home/rstudio/mnt_data/"
  out: "/home/rstudio/mnt_out/MES_project/"
  data: "/home/rstudio/mnt_data/"
  nb_tumors: "/home/rstudio/mnt_data/nb_tumors/"
  GO_db: "/home/rstudio/mnt_resources/GSEA_GeneSetEnrichments/"
  markers: "/home/rstudio/mnt_resources/MES_ADRN/"
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      tidy = TRUE, 
                      error = FALSE, 
                      messages = FALSE,
                      warning = FALSE)

library(parallel)
future::plan(strategy = 'future::multicore', workers = 20)
options(future.globals.maxSize = 8 * 10^9)

print(params)

source("/home/rstudio/MES_project/MES_project_setup.R")
```

```{r R packages: load}
#install.packages("Matrix", repos = "http://R-Forge.R-project.org")
suppressPackageStartupMessages(library("Seurat"))
suppressPackageStartupMessages(library("SeuratWrappers"))
suppressPackageStartupMessages(library("SeuratData"))
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("scCustomize"))
suppressPackageStartupMessages(library("RColorBrewer"))
suppressPackageStartupMessages(library("SCpubr"))
suppressPackageStartupMessages(library("ggVennDiagram"))
suppressPackageStartupMessages(library("statmod"))
suppressPackageStartupMessages(library("miloR"))
suppressPackageStartupMessages(library("SingleCellExperiment"))
suppressPackageStartupMessages(library("scater"))
suppressPackageStartupMessages(library("scran"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("patchwork"))
```

# ___Differential abundance testing for tumor clusters___


```{r}
NB_HR_patients_complete <- readRDS(file = 
                               paste0(params$nb_tumors,
                                      "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR_tumor_cluster.Rds"))
```

```{r eval=FALSE, include=FALSE}
NB_HR_patients_complete@meta.data$sample_name <- factor(NB_HR_patients_complete@meta.data$orig.ident)
NB_HR_patients_complete@meta.data$Treatment_Status <- factor(NB_HR_patients_complete@meta.data$Treatment_Status)

# add short treatment groups

Idents(NB_HR_patients_complete) <- NB_HR_patients_complete@meta.data$Treatment_Status 

NB_HR_patients_complete <- RenameIdents(NB_HR_patients_complete, 
                     "no treatment" = "untreated",
                     "sample collected after several courses of induction chemotherapy" = "treated")

NB_HR_patients_complete$treatment_groups <- NB_HR_patients_complete@active.ident


# add detailed treatment comparison groups

Idents(NB_HR_patients_complete) <- NB_HR_patients_complete@meta.data$anti_GD2 
NB_HR_patients_complete <- RenameIdents(NB_HR_patients_complete, 
                     "NA" = "at_diagnosis",
                     "No" = "after_chemo",
                     "Yes" = "after_chemo_immuno")


NB_HR_patients_complete$treatment_groups_detailed <- NB_HR_patients_complete@active.ident
table(NB_HR_patients_complete@meta.data$treatment_groups_detailed)

# add short MYCN status comparisons [so i can add MYCN as a covariate]

Idents(NB_HR_patients_complete) <- NB_HR_patients_complete@meta.data$MYCN

NB_HR_patients_complete <- RenameIdents(NB_HR_patients_complete, 
                                        "non-amplified" = "wt")

table(NB_HR_patients_complete@meta.data$MYCN)

NB_HR_patients_complete$MYCN_groups <- NB_HR_patients_complete@active.ident

table(NB_HR_patients_complete$MYCN_groups, 
      NB_HR_patients_complete$treatment_groups)

saveRDS(NB_HR_patients_complete, 
        file =  paste0(params$nb_tumors,
                  "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR_tumor_cluster.Rds"))
```

```{r eval=FALSE, fig.height=8, fig.width=15, include=FALSE}
Idents(NB_HR_patients_complete) <- NB_HR_patients_complete$liger_clusters

DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST, 
        raster = TRUE,
        shuffle = FALSE,
        label.size = 40,
        pt.size = 2)

DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST, 
        raster = FALSE,
        shuffle = FALSE,
        label.size = 40, 
        split.by = "treatment_groups")

Idents(NB_HR_patients_complete) <- NB_HR_patients_complete$treatment_groups

DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        raster = FALSE,
        shuffle = FALSE,
        label.size = 40, 
        split.by = "INRGSS_Stage") 

# only one stage L and one stage Ms high-risk NB patient, both untreated
# i think we should remove them as this will only hinder the analysis
```

```{r eval=FALSE, include=FALSE}
# Excluding stage Ms and L from the data
filter_by_stage <- NB_HR_patients_complete$INRGSS_Stage == "M"
NB_HR_patients_complete <- NB_HR_patients_complete[,filter_by_stage]

NB_HR_patients_complete$sample_name <- droplevels(NB_HR_patients_complete$sample_name)

NB_HR_patients_complete$sample_name <- factor(NB_HR_patients_complete$sample_name)
NB_HR_patients_complete$MYCN <- factor(NB_HR_patients_complete$MYCN)

saveRDS(NB_HR_patients_complete, 
        file = paste0(params$nb_tumors, "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR-M_tumor_cluster.Rds"))
```

```{r eval=FALSE, include=FALSE}
NB_HR_patients_complete <- readRDS(file = 
                               paste0(params$nb_tumors,
                                      "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR-M_tumor_cluster.Rds")) # stage M only
```

```{r fig.height=8, fig.width=15}
# Tumor cell clusters plot
Idents(NB_HR_patients_complete) <- NB_HR_patients_complete$liger_clusters

pdf(paste0(params$out,"figures/patel2024_UMAP_treatment_groups.pdf"),
    height=4, width=8)
DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        cols = COLOR_CODE_CLUST, 
        raster = FALSE,
        shuffle = FALSE,
        label.size = 60,
        pt.size = 0.5,
        split.by = "treatment_groups")
dev.off()

table(NB_HR_patients_complete$treatment_groups)

# MYCN status plot
Idents(NB_HR_patients_complete) <- NB_HR_patients_complete$treatment_groups

pdf(paste0(params$out,"figures/patel2024_UMAP_MYCN_groups.pdf"),
    height=4, width=8)
DimPlot(NB_HR_patients_complete, 
        seed = 2805, 
        raster = FALSE,
        shuffle = FALSE,
        label.size = 60, 
        pt.size = 0.5,
        split.by = "MYCN_groups") 
dev.off()

table(NB_HR_patients_complete$MYCN_groups)
```

```{r eval=FALSE, include=FALSE}
filter_by_MYCN_amp <- NB_HR_patients_complete$MYCN == "amplified"
NB_HR_patients_complete_stage_M_MYCN_amp <- NB_HR_patients_complete[,filter_by_MYCN_amp]

filter_by_MYCN_wt <- NB_HR_patients_complete$MYCN == "wt"
NB_HR_patients_complete_stage_M_MYCN_wt <- NB_HR_patients_complete[,filter_by_MYCN_wt]
```

```{r eval=FALSE, include=FALSE}
saveRDS(NB_HR_patients_complete_stage_M_MYCN_amp, 
        file = paste0(params$nb_tumors,
                      "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR-M-MYCNamp_tumor_cluster.Rds"))

saveRDS(NB_HR_patients_complete_stage_M_MYCN_wt, 
        file = paste0(params$nb_tumors,
                      "patel2024.scRNA/data/scRNA-seq/Rds/patel2024_NB-HR-M-MYCNwt_tumor_cluster.Rds"))
```

# ___MiloR DA analysis___

```{r miloR - prepare data}
# MiloR doesn't use the "original" clusters but rather creates neighborhoods
# based on similar logFC (note the fold change here is the number of cells !!
# per cell type)

# 1. pre-processing
Idents(NB_HR_patients_complete) <- NB_HR_patients_complete$liger_clusters

NB_HR_patients_complete <- RunPCA(NB_HR_patients_complete,  
                            npcs = 50,
                            reduction.name = "PCA")

# 2. save as single cell experiment (SCE)
rna_milo <- Milo(as.SingleCellExperiment(NB_HR_patients_complete))
rna_milo

# 3. add log counts
#logcounts(rna_milo) <- log(counts(rna_milo) + 1)

# 4. add PCA and UMAP reductions to the SCE 
reducedDim(rna_milo, "PCA") <- reducedDim(as.SingleCellExperiment(NB_HR_patients_complete), 
                                           "PCA")
reducedDim(rna_milo, "UMAP") <- reducedDim(as.SingleCellExperiment(NB_HR_patients_complete), 
                                           "UMAP")
# 5. QC / sanity check
plotUMAP(rna_milo, 
         colour_by="treatment_groups") + 
  plotUMAP(rna_milo, 
           colour_by="MYCN")

# 6. Constructing kNN graph   
rna_milo <- buildGraph(rna_milo,  
                       k = 10,  
                       d = 30)

# MiloR tool defined a neighbourhood of an index cell as the group of cells 
# connected by an edge in the KNN graph to the index cell. For efficiency, 
# testing for DA in the neighbourhood is not perfomed for every cell, but on 
# a subset of representative cells, using a KNN sampling algorithm used by 
# Gut et al. 2015.

# 7. Define cell neighbourhoods, with k = 10
rna_milo <- makeNhoods(rna_milo, 
                       prop = 0.1, 
                       k = 10,  
                       d = 30, 
                       refined = TRUE
                       ) 
#### refinement_scheme = "graph" -> Make nhoods using graph-only FASTER

plotNhoodSizeHist(rna_milo)

# 8. Counting cells per neighborhood
rna_milo <- countCells(rna_milo, 
                       meta.data = data.frame(colData(rna_milo)),
                       samples="sample_name") # make an "nhood x patients" counts matrix

# 9. Defining experimental design
rna_design <- data.frame(colData(rna_milo))[,c("sample_name", 
                                               "treatment_groups",
                                               "MYCN_groups")]
rna_design <- distinct(rna_design)
rownames(rna_design) <- rna_design$sample_name
rna_design

# 10. Computing neighborhood connectivity [used to compute p-values]
rna_milo <- calcNhoodDistance(rna_milo, 
                              d = 30)
```
## ___Untreated vs. treated HR-NB primary tumors___

```{r miloR - DA testing v1}

comparisons <- c("treatment_groupsuntreated - treatment_groupstreated")

# DA testing
da_results <- testNhoods(rna_milo, 
                         design = ~ treatment_groups + MYCN_groups,  
                         design.df = rna_design)
                      #  model.contrasts = comparisons
head(da_results)

da_results %>%
  arrange(- SpatialFDR) %>%
  head() 

rna_milo <- buildNhoodGraph(rna_milo)

# QUALITY CHECK OF THE DIFFERENTIAL ABUNDANCE TESTING

# a. Neighborhoods plot
nh_graph_pl <- plotNhoodGraphDA(rna_milo, 
                                da_results, 
                                layout = "UMAP", 
                                alpha = 0.05) 

# b. Single-cell UMAP
umap_pl <- plotReducedDim(rna_milo, 
                          dimred = "UMAP", 
                          colour_by = "sample_name",
                          text_by = "treatment_groups", 
                          text_size = 3, 
                          point_size = 0.5) +
  guides(fill = "none")

# c. combined plots
umap_pl + nh_graph_pl +
  plot_layout(guides = "collect")

# d.  logFC vulcano plot [to assess where to set the cut-offs]
ggplot(da_results, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1.3,colour = "darkgreen",linewidth = 1) + 
  geom_hline(yintercept = 1, colour = "lightgreen",linewidth = 1) + 
  geom_vline(xintercept = c(-3,3),colour = "lightblue",linewidth = 1) + 
  geom_vline(xintercept = c(-1.5,1.5),colour = "darkblue",linewidth = 1) 
```

## ___Downstream analysis___

```{r miloR - results}
table(da_results$SpatialFDR < 0.1)
table(abs(da_results$logFC) > 1.5)

# annotate hood with original cell type annotation
da_results <- annotateNhoods(rna_milo, 
                             da_results, 
                             coldata_col = "liger_clusters")

# define cut-off for homogeneous vs. heterogeneous neighborhoods
ggplot(da_results, aes(liger_clusters_fraction)) + geom_histogram(bins=200)

# remove hoods that contain mixed cell type populations (homogeneity above 75%)
da_results$cluster_fraction <- 
  ifelse(da_results$liger_clusters_fraction < 0.75, 
         "Mixed", da_results$liger_clusters)

# Automatic grouping of neighborhoods
da_results <- groupNhoods(rna_milo, 
                          da_results, 
                          da.fdr = 0.10,
                          max.lfc.delta = 3)
head(da_results)

plotNhoodGroups(rna_milo, da_results, layout = "UMAP")

# filter only significant populations (less than 5% false discovery rate and 
# fold change bigger than 3)
results <- da_results %>% 
  arrange(FDR) %>% filter(FDR < 0.1 & abs(logFC) >= 1.5) # based on the histogram

results_increase_after_treatment <- da_results %>% 
  arrange(FDR) %>% filter(FDR < 0.1 & logFC >= 1)

results_decrease_after_treatment <- da_results %>% 
  arrange(FDR) %>% filter(FDR < 0.1 & logFC <= -1)

plotDAbeeswarm(results, group.by = "liger_clusters")
plotDAbeeswarm(results, group.by = "NhoodGroup")

saveRDS(da_results, file = paste0(params$out,"Rds/milor_final_analysis.Rds"))
```

```{r DA analysis - MES cluster 4}
da_results <- readRDS(file = paste0(params$out,"Rds/milor_final_analysis.Rds"))
cluster_MES_filter <- da_results$cluster_fraction == "4"

cluster_MES_DA <- da_results[cluster_MES_filter,]

cluster_MES_DA_table <- data.frame(cluster_MES_DA$Nhood,
                                   cluster_MES_DA$logFC)

# Create density plot
pdf(paste0(params$out,"figures/patel2024_miloR_denstiyplot_MES_cluster4.pdf"),
    height = 4, width = 8)
ggplot(cluster_MES_DA_table, aes(x = cluster_MES_DA.logFC)) +
  geom_density(fill = "#961E1E", alpha = 0.5) +  # Density plot with shaded area
  theme_minimal() +
  labs(
    title = "DA of MES cluster 4 after treatment",
    x = "logFC",
    y = "Density"
  ) +
  theme(
    text = element_text(size = 14),
    plot.title = element_text(hjust = 0.5)
  )
dev.off()
```

```{r DA analysis - MES cluster 6}
cluster_MES_filter <- da_results$cluster_fraction == "6"

cluster_MES_DA <- da_results[cluster_MES_filter,]

cluster_MES_DA_table <- data.frame(cluster_MES_DA$Nhood,
                                   cluster_MES_DA$logFC)

# Create density plot
pdf(paste0(params$out,"figures/patel2024_miloR_denstiyplot_MES_cluster6.pdf"),
    height = 4, width = 8)
ggplot(cluster_MES_DA_table, aes(x = cluster_MES_DA.logFC)) +
  geom_density(fill = "#B94A4A", alpha = 0.5) + 
  theme_minimal() +
  labs(
    title = "DA of MES cluster 6 after treatment",
    x = "logFC",
    y = "Density"
  ) +
  theme(
    text = element_text(size = 14),
    plot.title = element_text(hjust = 0.5)
  )
dev.off()
```

```{r DA analysis - MES cluster 4+6+11}
cluster_MES_filter <- da_results$cluster_fraction == c("4","6","11")

cluster_MES_DA <- da_results[cluster_MES_filter,]

cluster_MES_DA$cluster_fraction <- as.factor(cluster_MES_DA$cluster_fraction)

cluster_MES_DA_table <- data.frame(cluster_MES_DA$cluster_fraction,
                                   cluster_MES_DA$logFC)

# Create density plot
pdf(paste0(params$out,"figures/patel2024_miloR_denstiyplot_MES_clusters.pdf"),
    height = 4, width = 12)
ggplot(cluster_MES_DA_table, aes(x = cluster_MES_DA.logFC,
                                 fill =  cluster_MES_DA.cluster_fraction)) +
  geom_density(alpha = 0.5) +  
  theme_minimal() +
   scale_fill_manual(values = c("#ff7b7b","#B94A4A","#961E1E")) +
  labs(
    title = "DA of all MES clusters after treatment",
    x = "logFC",
    y = "Density"
  ) +
  theme(
    text = element_text(size = 14),
    plot.title = element_text(hjust = 0.5)
  )
dev.off()
```

```{r DA analysis - ADR clusters}
cluster_ADR_filter <- da_results$cluster_fraction != c("4","6","11")

cluster_ADR_DA <- da_results[cluster_ADR_filter,]

cluster_ADR_DA$cluster_fraction <- as.factor(cluster_ADR_DA$cluster_fraction)

cluster_ADR_DA_table <- data.frame(cluster_ADR_DA$Nhood,
                                   cluster_ADR_DA$logFC)

# Sort by the first column
cluster_ADR_DA_table_sorted <- cluster_ADR_DA_table[order(cluster_ADR_DA_table[, 1]), ]
cluster_ADR_DA_table_sorted_unique <- unique(cluster_ADR_DA_table_sorted)

# Create density plot
pdf(paste0(params$out,"figures/patel2024_miloR_denstiyplot_ADR_clusters.pdf"),
    height = 4, width = 8)
ggplot(cluster_ADR_DA_table_sorted_unique, aes(x = cluster_ADR_DA.logFC)) +
  geom_density(alpha = 0.5) +  
  geom_density(fill = "#0055b3", alpha = 0.5) +
  theme_minimal()  +
  labs(
    title = "DA of all ADR clusters after treatment",
    x = "logFC",
    y = "Density"
  ) +
  theme(
    text = element_text(size = 14),
    plot.title = element_text(hjust = 0.5)
  )
dev.off()
```

```{r DA analysis - all cluster}
da_results$cluster_fraction <- as.factor(da_results$cluster_fraction)

cluster_DA_table <- data.frame(da_results$cluster_fraction,
                                   da_results$logFC)

colnames(cluster_DA_table) <- c("clusters", "logFC")

clust_order <- c("1", "2", "3", "4","5","6","7","8","9","10","11","12","Mixed")  
cluster_DA_table$clusters <- factor(cluster_DA_table$clusters, levels = clust_order)

# Create density plot 
pdf(paste0(params$out,"figures/patel2024_miloR_densityplot_all_clusters.pdf"),
    height = 5, width = 8)
ggplot(cluster_DA_table, aes(x = logFC,
                             fill = clusters)) +
  geom_density(alpha = 0.5) +  
  theme_minimal() +
   scale_fill_manual(values = COLOR_CODE_CLUST_v2) +
  labs(
    title = "DA of all clusters after treatment",
    x = "logFC",
    y = "Density"
  ) +
  theme(
    text = element_text(size = 14),
    plot.title = element_text(hjust = 0.5)
  )  + facet_wrap(~ clusters)
dev.off()
```

```{r}
COLOR_CODE_CLUST <- c("9" = "green", 
                      "4" = "blue", 
                      "5" = "purple")

DimPlot(NB_HR_patients_complete,
seed = 2805,
cols = COLOR_CODE_CLUST,
raster = FALSE,
shuffle = FALSE,
label.size = 40,
split.by = "treatment_groups",
pt.size = 0.25)
```

```{r}
# Extract metadata from the Seurat object
metadata <- NB_HR_patients_complete@meta.data

# Summarize read counts per cluster and treatment group
read_count_summary <- metadata %>%
  group_by(liger_clusters, treatment_groups) %>%
  summarise(mean_read = mean(nCount_RNA, na.rm = TRUE),
            median_read = median(nCount_RNA, na.rm = TRUE),
            .groups = "drop")

# Print the summary table
print(read_count_summary)

# Create a boxplot to visualize the distribution of read counts per cluster, split by treatment
ggplot(metadata, aes(x = factor(liger_clusters), y = nCount_RNA, fill = treatment_groups)) +
  geom_boxplot() +
  labs(title = "QC: Read Count per Cluster by Treatment Group",
       x = "Cluster",
       y = "log10 readCount (nCount_RNA)") +
  theme_minimal() +
  scale_y_log10()  # Optional: apply a log10 scale if read counts vary widely
```


```{r save results, eval=FALSE, include=FALSE}
write.table(results, 
            file = 
            paste0(params$out, "tables/patel2024_NB-HR_DA_testing_treatedVSuntreated.csv"),
                   sep = "\t", quote = FALSE)

saveRDS(rna_milo, 
        file = 
          paste0(params$out,
                      "Rds/patel2024_NB-HR_tumorCells_DAtesting.Rds"))
```


# ___nhood markers___

Find hood markers, than aggregating the neighbourhood expression profiles by experimental samples (the same used for DA testing), by setting aggregate.samples=TRUE. This way single-cells will not be considered as “replicates” during DGE testing, and dispersion will be estimated between true biological replicates. 

```{r eval=FALSE, include=FALSE}

da_nhood_markers <- findNhoodGroupMarkers(rna_milo, 
                                          da_results, 
                                          subset.row = rownames(rna_milo),
                                          aggregate.samples = TRUE, 
                                          sample_col = "sample_name")

# CLUSTER 4 DEGs between treated and untreated tumor samples

ggplot(da_nhood_markers, 
       aes(logFC_4, -log10(adj.P.Val_4))) + 
  geom_point(alpha=0.5, size=0.5) +
  geom_hline(yintercept = 3)

set.seed(42)
clust4_markers <- da_nhood_markers$GeneID[da_nhood_markers$adj.P.Val_4 < 0.1 & 
                                     abs(da_nhood_markers$logFC_4) > 0.1]

plotNhoodExpressionGroups(rna_milo, 
                          da_results, 
                          features=intersect(rownames(rna_milo), clust4_markers),
                          subset.nhoods = da_results$NhoodGroup %in% '4', 
                          scale=TRUE,
                          grid.space = "fixed")

# DGE testing within a group
DGE <- testDiffExp(rna_milo, 
                          da_results, 
                          design = ~ treatment_groups, 
                          meta.data = data.frame(colData(rna_milo)),
                          subset.row = rownames(rna_milo))
                          #subset.nhoods=da_results$NhoodGroup=="4")

signif_genes <- DGE %>% arrange(FDR) %>% filter(FDR < 0.01 & abs(logFC) >= 0.1)
signif_genes
```


# ___DA testing with super-FACS method___

source: https://bioconductor.org/books/3.13/OSCA.multisample/differential-abundance.html

```{r}
sce <- as.SingleCellExperiment(NB_HR_patients_complete)
abundances <- table(sce$liger_clusters, 
                    sce$sample_name) 
abundances <- unclass(abundances) 
head(abundances)

library(edgeR)

metadata.info <- colData(sce)[match(colnames(abundances), 
                                    sce$sample_name),]
dge <- DGEList(abundances, 
               samples = metadata.info)
dge 

# For a DA analysis of cluster abundances, filtering is generally not required as most clusters will not be of low-abundance (otherwise there would not have been enough evidence to define the cluster in the first place).

#keep <- filterByExpr(dge, group = dge$sample_name$treatment_groups)
#dge <- dge[keep,]
#summary(keep)

design <- model.matrix(~ treatment_groups + MYCN_groups,
                       dge$sample) # dge$sample is in this case the metadata table

dge <- estimateDisp(dge, design, 
                    trend = "none")
summary(dge$common.dispersion)

plotBCV(dge, 
        cex = 1)

fit.ab <- glmQLFit(dge, 
                   design, 
                   robust = TRUE, 
                   abundance.trend = FALSE)
summary(fit.ab)
summary(fit.ab$df.prior)

plotQLDisp(fit.ab, 
           cex = 1)

res <- glmQLFTest(fit.ab, 
                  coef = ncol(design))
summary(decideTests(res))
topTags(res)
```

```{r}
# alternative hypothesis: we assume most cell types will not change in 
# abundance after treatment

dge2 <- calcNormFactors(dge)
dge2$sample_name$norm.factors

dge2 <- estimateDisp(dge2, 
                     design, 
                     trend = "none")
fit.ab2 <- glmQLFit(dge2, 
                    design, 
                    robust = TRUE, 
                    abundance.trend = FALSE)
res2 <- glmQLFTest(fit.ab2, 
                   coef = ncol(design))
topTags(res2, n = 10)

summary(decideTests(res2))
topTags(res2)
```

# ___Alternative DA testing method___

```{r eval=FALSE, include=FALSE}
library(DAseq)
python2use <- "/home/rstudio/.virtualenvs/r-reticulate/bin/python"

# 20 patients with NB histology, high-risk tumors and information on treatment
NB_HR_patients_complete <- readRDS(file = paste0(params$out,
                                  "patel2024/Rds/patel2024_NB-HR_tumor_cluster.Rds"))

filter_by_treatment <- NB_HR_patients_complete$Treatment_Status  == "no treatment"
NB_HR_patients_complete <- NB_HR_patients_complete[,filter_by_treatment]
table(NB_HR_patients_complete$Treatment_Status)


# Create a new column based on the existing column
NB_HR_patients_complete$treatment_groups <- ifelse(NB_HR_patients_complete$Treatment_Status == "no treatment", "untreated", "treated")

# Verify the new column has been added
table(NB_HR_patients_complete$treatment_groups)

labels_treat <- X.label.info[X.label.info$condition == "treated", "treatment_groups"]
labels_untrea <- X.label.info[X.label.info$condition == "untreated", "treatment_groups"]

da_cells <- getDAcells(  
    X = NB_HR_patients_complete,
    cell.labels = NB_HR_patients_complete$treatment_groups,
    labels.1 = labels_treat,
    labels.2 = labels_untrea,
    k.vector = seq(50, 500, 50))
```



# ___SessionInfo___

```{r}
sessionInfo()
```

